---
title: "DiSTATIS"
output:
  word_document: default
  html_document: default
---

```{r DiSTATIS_setup, include=FALSE}
rm(list = ls())
graphics.off()

knitr::opts_chunk$set(echo = TRUE)

#install.packages('DistatisR')

library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)

library(scales)

library(TInPosition)
library(officer)
library(flextable)
library(rvg)

library(useful)
library(DistatisR)



```


# DiSTATIS



## Data
We have the data related 18 wines, that have been assesed by 19 different experts.
Wines tested basically are of 3 colors (Red, White, Rose).

TO find a general trend amongst the experts I randomly assign them gender (this was not given in data, any conclusion from this will not be significant).
```{r}
Raw <- read.csv('19ExpertsFreeSort18Wines3Colors.csv', row.names=3)
Sort <- Raw[,-c(1:2)]

BoysGirls = c('f','m','f','f','m','m','m','m','f','m','m','f','f','m','m','m','m','f','m')
Judges <- paste0(BoysGirls,1:length(BoysGirls))
head(Sort)
#dim(Sort)
```


### Splitting into cubes based on the assessors (J1,J2,.. etc)
Each assessor will have a contingency table sort for the rows, where in the 2 or more rows having matching values will be 0,
rest all will be 1.
Similarly there will be cubes for all assessors.

### Distance Matrices
```{r}
#---------------------------------------------------------------------
# 2. Create the set of distance matrices
#  (one distance matrix per assessor)
#    (use the function DistanceFromSort)
DistanceCube <- DistatisR::DistanceFromSort(Sort)
```


## Running DiSTATIS
```{r}
#---------------------------------------------------------------------
# 3. Call the DISTATIS routine with the cube of distance as parameter
testDistatis <- DistatisR::distatis(DistanceCube)
# The factor scores for the beers are in
# testDistatis$res4Splus$F
# the factor scores for the assessors are in (RV matrice)
#  testDistatis$res4Cmat$G
```

### Bootstrap
```{r}
#---------------------------------------------------------------------
# 4. Inferences on the beers obtained via bootstrap
# 4.1 Get the bootstrap factor scores (with default 1000 iterations)
BootF <- BootFactorScores(testDistatis$res4Splus$PartialF)
#
#---------------------------------------------------------------------
#*********************************************************************
# End of computations
# We have now in testDistatis
#1 "$res4Cmat"  "Results from the C matrix (see notation)"
#2 "$res4Splus" "Results from the S+ matrix (see notation)"
#

```



### Scree Plot
```{r}
#---------------------------------------------------------------------
# 5. The graphs
#  5.1 First the C matrix
# Look at the scree to start with
#
ev4C <- testDistatis$res4Cmat$eigValues
Scree.1 <-  PlotScree(ev = ev4C,
          p.ev = NULL, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F",
          title = "RV-mat: Explained Variance per Dimension",plotKaiser = TRUE)
zeScree.Rv <- recordPlot()
```

```{r echo=FALSE}
# Plot the assessor matrix
G <- testDistatis$res4Cmat$G
col4B <- 'blue'
col4G <- 'magenta'
col4Judges <- rep(col4G,length(BoysGirls))
col4Judges[BoysGirls == 'm'] <- col4B


Eigs <- testDistatis$res4Cmat$eigValues
tau <- testDistatis$res4Cmat$tau

#Creating labels for Dim 1 & 2
expert_labels_12 <- createxyLabels.gen(1,2, lambda = Eigs,tau = tau)

```


### Graph for the assessors

**Group Means and CI cant be plotted for the assessors as they cannot be differentiated on any variables**
```{r}
#---------------------------------------------------------------------
#---------------------------------------------------------------------
# A graph for the Judges set
baseMap.j <- PTCA4CATA::createFactorMap(G,
                                        title = 'The Rv map',
                                        col.points   = col4Judges,
                                        alpha.points =  .3,
                                        col.labels = col4Judges
                                        )
# A graph for the J-set
aggMap.j <- baseMap.j$zeMap_background + # background layer
           baseMap.j$zeMap_dots + baseMap.j$zeMap_text + expert_labels_12
          
# We print this Map with the following code
#dev.new()
print(aggMap.j)

```

### Hull for the experts plot
```{r}
GraphTJ.Hull.100 <- MakeToleranceIntervals(G,
                                  as.factor(BoysGirls),
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = unique(col4Judges),
                                          line.size = .5,
                                          line.type = 3,
                                          alpha.ellipse = .1,
                                          alpha.line = .4,
                                          p.level = 1, # full Hulls
                                          type = 'hull' #
                                          # use 'hull' for convex hull
)
#---------------------------------------------------------------------
# Create the map
aggMap.j.withHull <- baseMap.j$zeMap_background + # background layer
                           baseMap.j$zeMap_dots + GraphTJ.Hull.100 + expert_labels_12

#dev.new()
print(aggMap.j.withHull)

```

### Mean
```{r}
#---------------------------------------------------------------------
# Plot for the group means computed on the factor scores
#---------------------------------------------------------------------
# First compute the means
JudgesMeans.tmp <- aggregate(G, list(BoysGirls), mean) # compute the means
JudgesMeans <- JudgesMeans.tmp[,2:ncol(JudgesMeans.tmp )] # drop var 1
rownames(JudgesMeans) <- JudgesMeans.tmp[,1] # use var 1 to name the groups
#---------------------------------------------------------------------
# a vector of color for the means
col4Means <- unique(col4Judges)
#---------------------------------------------------------------------
# create the map for the means
MapGroup    <- PTCA4CATA::createFactorMap(JudgesMeans,
                                  axis1 = 1, axis2 = 2,
                                  constraints = baseMap.j$constraints,
                                          title = NULL,
                                          col.points = col4Means,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4Means,
                                          text.cex = 4,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                          col.background = adjustcolor("lavender",
                                             alpha.f = 0.2),
                                          force = 1, segment.size = 0)
# The map with observations and group means
aggMap.j.withMeans <- aggMap.j.withHull +
                    MapGroup$zeMap_dots + MapGroup$zeMap_text 

#dev.new()
print(aggMap.j.withMeans)

```

### For Rows

```{r}
# Go for the beers now
# First we fix a bit of shamefull absentmindness:
#   The eigenvalues of the compromise matrix are not available
#   in DistatisR.
#   So we recompute them here
ev4S <- eigen(testDistatis$res4Splus$Splus,
                symmetric = TRUE, only.values = TRUE)$values
# A scree for the compromise
Scree.S <-  PlotScree(ev = ev4S,
                   p.ev = NULL, max.ev = NULL, alpha = 0.05,
                   col.ns = "#006D2C", col.sig = "#54278F",
                   title = "S-mat: Explained Variance per Dimension",plotKaiser = TRUE)
zeScree.S <- recordPlot()
```

```{r}
Fi <- testDistatis$res4Splus$F
col4Wines <- prettyGraphsColorSelection(nrow(Fi))
```

### Graphs for I set
```{r}
# Graphs for the I set
#---------------------------------------------------------------------
# Create the base map
constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
baseMap.i <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = col4Wines,
                                        col.labels = col4Wines,
                        constraints = constraints4Fi,
                                        alpha.points =  .4)
#---------------------------------------------------------------------
# We are  interested about the labels here
# so we will use dots and labels
#---------------------------------------------------------------------
# Plain map with color for the I-set
aggMap.i <- baseMap.i$zeMap_background + baseMap.i$zeMap_dots +
                                         baseMap.i$zeMap_text + expert_labels_12
#---------------------------------------------------------------------
# print this Map
#dev.new()
print(aggMap.i)
```

### Confidence Interval
```{r}
#---------------------------------------------------------------------
# Create Confidence Interval Plots
# use function MakeCIEllipses from package PTCA4CATA
#
constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
GraphElli <- MakeCIEllipses(BootF[,1:2,],
                      names.of.factors = c("Factor 1","Factor 2"),
                      alpha.line = .5,
                      alpha.ellipse = .3,
                      line.size = .5,
                      line.type = 3,
                      col = col4Wines,
                      p.level = .95 )
#---------------------------------------------------------------------
# create the I-map with Observations and their confidence intervals
#
aggMap.i.withCI <-  aggMap.i +  GraphElli + MapGroup$zeMap_text + expert_labels_12
#---------------------------------------------------------------------
# plot it!
#dev.new()
print(aggMap.i.withCI)
```

### Partial Factor Scores

```{r}
# Old graph with links to partial factor scores
# Not that informative for sorting tasks
# Change names of the assessors
partF <- testDistatis$res4Splus$PartialF
dimnames(partF)[[3]] <- as.character(1:dim(partF)[3])
PartialF <- GraphDistatisPartial(FS = testDistatis$res4Splus$F,
                     PartialFS = partF,
                     axis1 = 1, axis2 = 2, constraints = NULL,
                     item.colors = col4Wines,
                     participant.colors = NULL,
                     ZeTitle = "Distatis-Partial",
                     Ctr=NULL, color.by.observations = TRUE,
                     nude = FALSE, lines = TRUE)
# save the graphs in F.and.PartialF
F.and.PartialF <- recordPlot()

```


```{r echo=FALSE}
# Some classification now
# First plain k-means
set.seed(42)
wines.kMeans <- kmeans(x = Fi , centers = 3)
#---------------------------------------------------------------------
# Now to get a map by cluster:
col4Clusters  <- createColorVectorsByDesign(
              makeNominalData(
              as.data.frame(wines.kMeans$cluster)  ))
#col4Clusters$oc = c('Red','Yellow','Pink')
#rownames(col4Clusters$gc) = c("wines.kMeans$cluster.2", "wines.kMeans$cluster.1", "wines.kMeans$cluster.3")
#col4Clusters$oc = c('Pink','Red','Yellow','Red','Pink','Pink','Red','Yellow','Red','Yellow','Pink','Yellow','Pink','Red','Pink','Yellow','Red','Yellow')
#=====================================================================
```


### Graph fo rI set
```{r}
#=====================================================================
#---------------------------------------------------------------------
# Graphs for the I set
#---------------------------------------------------------------------
# Create the base map
# constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
baseMap.i.km <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = col4Clusters$oc,
                                        col.labels = col4Clusters$oc,
                                        constraints = constraints4Fi,
                                        alpha.points =  .4)
#---------------------------------------------------------------------
# We are  interested about the labels here
# so we will use dots and labels
#---------------------------------------------------------------------
# Plain map with color for the I-set
aggMap.i.km <- baseMap.i.km$zeMap_background +
  baseMap.i.km$zeMap_dots + baseMap.i.km$zeMap_text
# print
#dev.new()
print(aggMap.i.km)
```

### Adding cluster names
```{r}
#---------------------------------------------------------------------
# Add the cluster names
#---------------------------------------------------------------------
# get the color order in the c=good order
col4C <- col4Clusters$gc[sort(rownames(col4Clusters$gc),
                                     index.return = TRUE)$ix]
# create the map for the means
map4Clusters    <- PTCA4CATA::createFactorMap(wines.kMeans$centers,
                                          axis1 = 1, axis2 = 2,
                                          constraints = constraints4Fi,
                                          title = NULL,
                                          col.points = col4C,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4C,
                                          text.cex = 6,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                                          col.background =
                          adjustcolor("lavender", alpha.f = 0.2),
                                          force = 1, segment.size = 0)
# The map with observations and group means
aggMap.i.withCenters <- aggMap.i.km  +
               map4Clusters$zeMap_dots + map4Clusters$zeMap_text
#
#dev.new()
print(aggMap.i.withCenters)
#---------------------------------------------------------------------
```

## Cluster analysis
```{r}
#*********************************************************************
# A cluster analysis
wine.hc <- hclust(d = dist(Fi),
                  method = 'ward.D2' )

plot.tree <- plot(wine.hc,  main = "Wines. Ward's method")
hc.tree <- recordPlot()
#dev.new()
print(hc.tree)

hc.3.cl <- rect.hclust(wine.hc, k = 3,
               border = c('darkorchid',
                          'darkolivegreen4','darkgoldenrod3')
                          )
hc.tree.3c <- recordPlot()
#dev.new()
print(hc.tree.3c)

```



### Optimal num of clusters
```{r}
# Optimal number of clustera
#  install.packages('useful')
best.wines <- useful::FitKMeans(Fi, max.clusters = 3,
                                seed = 314)
print(best.wines) # when Hartigan parameter > 10 => add a cluster
#dev.new()
plot.harti <- useful::PlotHartigan(best.wines)
print(plot.harti)
# here effect of small N. go for 4 clusters ?
```

## Conclusion

