
---
title: "PCA on IBM Data"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "Spetember 22, 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r PCA_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())

devtools::install_github('HerveAbdi/PTCA4CATA')
#devtools::install_github('HerveAbdi/DistatisR')
devtools::install_github('HerveAbdi/data4PCCAR')

library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)
```

# Principal Component Analysis

Principal Component Analysis(PCA) is used to analyze the continuous variables in a given dataset, which helps us in interpreting which all variables contribute more to the flow of data and in the end cluster the data in various groups.

Consider a 2D dataset(has 2 variables/columns) that has all its observations(rows) plotted on the X-Y axis.

We do a little preprocessing before we begin, i.e. we center and scale the data.
Centering: Refers to subtracting mean of each column from each of its points.
Scaling: Normalization

Now we plot the new point on the XY axis.

We try to find a line among all these points such that **minimizes** sum of square of perpendicular distance each point from that line.

**OR**
Try to find a line that **maximizes** the distance of projection of each point onto this line from the origin.

This line that we get is the **Principal Component 1(PC1)**. It will have the maximum inertia(variance), or it kind of pulls the majority of the data.

Next, **Principal Component 2(PC2)** is a line orthogonal to PC1(*We take the orthogonal line as next component in order to reduce any effect from the PC1, so we will just get the effect of PC2 on the observations*)

**Factor Score**: Instead of distance from XY axis, we take the distance w.r.t PC1 and PC2 thereby giving Factor scores of the observations.
                  PCA gives one map of factor scores, from which we can make out the observations that have more effect on the components.
                  For intrepretation in this graph we consider the distances between factor score.

**Loadings**: These corresspond to the variables/columns being plotted on the PC axis.
              Another PCA map gives loadings of variables.
              For intrepretation we consider the angle between various variables.

Based on boh these maps we find the important variables and factors that drive the overall data, adn divide it into various groups.



## Data set: IBM-HR-Emplyee-WithAttrition

The dataset given is regarding IBM employees, available to the HR section of the company to take decisions regarding attrition.

It consists of data of 237 employees (rows), with 32 variables(columns) defining each employee based on position,education,department etc.


```{r PCA_data_set, include=FALSE}
ibm.emp.data <- read.xlsx("IBM-HR-Emplyee-WithAttrition.xlsx", sheetName = "Sheet1",header=TRUE)
rownames(ibm.emp.data) <- ibm.emp.data$Subj
ibm.emp.data <- ibm.emp.data[-c(1)]
cols <- c('Age','Gender','JobRole','MonthlyIncome','TotalWorkingYears','JobSatisfaction')
head(ibm.emp.data[cols])


```

Of the 32 variables, we have following:
*Quantitative Variables* :Subj, Age, Monthly Income, DailyRate, HourlyRate, MonthlyRate,PercentSalaryHike etc
*Nominal Variables*: BusinessTravel, Department, EducationField,Attrition,JobRole
*Ordinal Variables*: PerformanceRating, JobLevel, JobInvolvement

### Corelation Plot
```{r echo=FALSE}
ibm.emp.data$Education <- factor(ibm.emp.data$Education)
#sapply(ibm.emp.data, class)
new.ibm.emp.data <- ibm.emp.data[sapply(ibm.emp.data, function(x) !is.factor(x))]

cor.res <- cor(new.ibm.emp.data)
corrplot(cor.res,method = c("ellipse"),type = 'lower')
```

From the correlation plot, we see that year related variables have significant positive corelation amongst themselves.
Apart from the above variabels we have +ve correlation between Age-Monthly Income, Job Level,Num of COmpanies Worked.
Also we see signinificant relation between Job Level-Monthly Income.


### Heat Map
```{r}
dev.new()
color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new.ibm.emp.data))

corrMatBurt.list <- phi2Mat4BurtTable(new.ibm.emp.data)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corr4PCA.r <- corrplot::corrplot(
as.matrix(corrMatBurt.list$phi2.mat),
method="color", col=col(200),
type="upper",
addCoef.col = "black", # Add coefficient of correlation
tl.col=color4Var,
tl.srt = 45, #Text label color and rotation
number.cex = .5,
diag = TRUE # needed to have the color of variables correct
)

```
After the removal of qualitative data, the heat map shows,strong +ve correlation between Monthly Rate and all other variables.
ALso we see similar trend with Monthly Income & Hourly Rate.
Perfect COrrelation between Performance Rating and Percent Salary Hike.



## Create DESIGN variables.

### 1st Design Variable : Job Level

We choose Job Level as 1st design variable.
Job Level is related to the position of employee in the company.
In this data we have 5 levels for Job Level (1,2,3,4,5)
With 1 being the position wise seniormost employess and 5 accordingly loewr ranked ones.
```{r PCA_DESIGN, include = FALSE}
#Create an empty list, called DESIGN
DESIGN <- list()

#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$JobLevel$labels <- unique(ibm.emp.data$JobLevel)


#Storing the categorical values in a vector
DESIGN$rows$JobLevel$vec <- ibm.emp.data$JobLevel


#Convert the vector to a matrix
DESIGN$rows$JobLevel$mat <- makeNominalData(as.matrix(DESIGN$rows$JobLevel$vec))

```


```{r PCA_colors_2, include= FALSE}

#As our group has 5 levels we specify same number of colors

DESIGN$rows$JobLevel$color_groups <- c("red", "green", "magenta", "blue","yellow")

#First, copy the group names
DESIGN$rows$JobLevel$color_observ <- as.matrix(DESIGN$rows$JobLevel$vec)

#Then, for each group, replace the group name with the group's color
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="1")]  <- DESIGN$rows$JobLevel$color_groups[1]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="2")]  <- DESIGN$rows$JobLevel$color_groups[2]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="3")]  <- DESIGN$rows$JobLevel$color_groups[3]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="4")]  <- DESIGN$rows$JobLevel$color_groups[4]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="5")]  <- DESIGN$rows$JobLevel$color_groups[5]

```


### 2nd Design Variable : Performance Rating
We choose Performance Rating as 2nd design variable.
Performance Rating is related to the year end evaluation of employees.
In this data we have 2 levels for Performance Rating (3,4)
With 4 being the best and 3 the next best.

```{r PCA_DESIGN_1,include=FALSE}
#Repeating the same process as above for Performance Rating
DESIGN_1 <- list()

#We take all the unique values possible in the specified column if the dataframe
DESIGN_1$rows$PerfRating$labels <- unique(ibm.emp.data$PerformanceRating) 

#Storing the categorical values in a vector
DESIGN_1$rows$PerfRating$vec <- ibm.emp.data$PerformanceRating

#Convert the vector to a matrix
DESIGN_1$rows$PerfRating$mat <- makeNominalData(as.matrix(DESIGN_1$rows$PerfRating$vec))
```



```{r PCA_colors_2_pr,include=FALSE}

#As our group has 2 levels we specify same number of colors

DESIGN_1$rows$PerfRating$color_groups <- c("green", "blue")

#First, copy the group names
DESIGN_1$rows$PerfRating$color_observ <- as.matrix(DESIGN_1$rows$PerfRating$vec) 

#Then, for each group, replace the group name with the group's color
DESIGN_1$rows$PerfRating$color_observ[which(DESIGN_1$rows$PerfRating$vec=="3")]  <- DESIGN_1$rows$PerfRating$color_groups[1]
DESIGN_1$rows$PerfRating$color_observ[which(DESIGN_1$rows$PerfRating$vec=="4")]  <- DESIGN_1$rows$PerfRating$color_groups[2]
```


## Running the PCA

We run the epPCA and epPCA.inference model by passing the data containing only quantitative variables and also the design variables.

```{r PCA_analyze, echo = TRUE}
new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(JobLevel))
new.ibm.emp.data2 <- subset(new.ibm.emp.data, select = -c(PerformanceRating))

res_pca_d1 <- epPCA(new.ibm.emp.data1, center = TRUE,DESIGN = DESIGN$rows$JobLevel$color_observ, scale = TRUE, graphs = TRUE)

res_pca_d2 <- epPCA(new.ibm.emp.data2, center = TRUE,DESIGN_1$rows$PerfRating$color_observ, scale = TRUE, graphs = TRUE)

# do inference PCA

res_pca.inf_d1 <- epPCA.inference.battery(new.ibm.emp.data1,scale = 'ss1', DESIGN = DESIGN$rows$JobLevel$color_observ, make_design_nominal = TRUE)

res_pca.inf_d2 <- epPCA.inference.battery(new.ibm.emp.data2,scale = 'ss1', DESIGN = DESIGN_1$rows$PerfRating$color_observ, make_design_nominal = TRUE)
```



### Scree Plot
We first do a scree plot with inference, which shows how many components are required to define the data, and which all out of these are the significant ones.  

- We find that 3 component come up as significant during permutation test.

```{r PCA_scree_plot}
#Get the Eigen Values list
EigenValues <- res_pca_d1$ExPosition.Data$eigs


PlotScree(ev = EigenValues, 
       p.ev =  res_pca.inf_d1$Inference.Data$components$p.vals,
       title = "Scree Plot with Permutation Test - PCA",
       plotKaiser = TRUE
          )
#dev.new()
a001a.screePlot <- recordPlot()
#ev.new()
print(a001a.screePlot)

```


```{r include=FALSE}
Fi <- res_pca_d1$ExPosition.Data$fi
Fi.1 <- res_pca_d2$ExPosition.Data$fi
Fj <- res_pca_d1$ExPosition.Data$fj
Fj.1 <- res_pca_d2$ExPosition.Data$fj
Cj <- res_pca_d1$ExPosition.Data$cj
Cj.1 <- res_pca_d2$ExPosition.Data$cj
Eigs <- res_pca_d1$ExPosition.Data$eigs
tau <- res_pca_d1$ExPosition.Data$t

color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new.ibm.emp.data1))
col4Levels <- data4PCCAR::coloringLevels(rownames(Fj), color4Var)
col4Labels <- col4Levels$color4Levels

varCtr <- data4PCCAR::ctr4Variables(Cj)



absCtrVar <- as.matrix(varCtr) %*% diag(Eigs)
varCtr12 <- (absCtrVar[,1] + absCtrVar[,2]) /(Eigs[1] + Eigs[2])
importantVar <- (varCtr12 >= 1 / length(varCtr12))
col4ImportantVar <- color4Var
col4NS <- 'gray90'
col4ImportantVar[!importantVar] <- col4NS
#Color for importance in  variables
col4Levels.imp <- data4PCCAR::coloringLevels(rownames(Fj),col4ImportantVar)


```


### Factor scores Based on Job Level
Factor scores are the coordinates of the 237 employees on the components. The distances between them show which employees are most similar. Factor scores (states) can be color-coded to help interpret the components.


#### Between Component 1 & 2:
```{r factor scores JOb Level 12,echo=FALSE}

#Plotting Factor Scores for Dim 1 & 2
baseMap.jl.12 <- createFactorMap(Fi, col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Factor scores by Job Level",display.labels = FALSE,
                             display.points = TRUE,axis1 = 1,axis2 = 2)  

#Creating labels for Dim 1 & 2
x_y_labels_1_2 <- createxyLabels.gen(1,2, lambda = Eigs,tau = tau)

#Calculating the means for Dim 1 & 2
data_means_JL <- PTCA4CATA::getMeans(Fi, ibm.emp.data$JobLevel)
rownames(data_means_JL) = c("JL5","JL4","JL3","JL2","JL1")


#Mapping the means
MapGroup_jl_12    <- PTCA4CATA::createFactorMap(data_means_JL,
                            # use the constraint from the main map
                            constraints = baseMap.jl.12$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,axis1 = 1,axis2 = 2,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)

#Adding all the plots
jl_1_2 <- baseMap.jl.12$zeMap + x_y_labels_1_2+MapGroup_jl_12$zeMap_dots + MapGroup_jl_12$zeMap_text

#dev.new()

print(jl_1_2)


```

#### Confidence Intervals
```{r}
# Confidence Intervals ----
# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(res_pca_d1$ExPosition.Data$fi, 
                                 design = DESIGN$rows$JobLevel$color_observ,
                                 niter = 100,
                                 suppressProgressBar = TRUE)


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            p.level = .95
)

JL.withCI <-  jl_1_2 +  GraphElli

#dev.new()
print(JL.withCI)

```


#### Between Component 2 & 3:
```{r factor scores JOb Level 23,echo=FALSE}
#Plotting Factor Scores for Dim 1 & 2
baseMap.jl.23 <- createFactorMap(Fi, col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Factor scores by Job Level",display.labels = FALSE,
                             display.points = TRUE,axis1 = 2,axis2 = 3)  

#Creating labels for Dim 2 & 3
x_y_labels_2_3 <- createxyLabels.gen(2,3, lambda = Eigs,tau = tau)

#Mapping the means
MapGroup_jl_23    <- PTCA4CATA::createFactorMap(data_means_JL,
                            # use the constraint from the main map
                            constraints = baseMap.jl.23$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,axis1 = 2,axis2 = 3,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)

#Adding all the plots
jl_2_3 <- baseMap.jl.23$zeMap + x_y_labels_2_3+MapGroup_jl_23$zeMap_dots + MapGroup_jl_23$zeMap_text
#dev.new()
print(jl_2_3)
```



### Factor scores Based on Performance Rating
We plot and color based on Performacne Rating

#### Between Component 1 & 2:
```{r}

#Plotting Factor Scores for Dim 1 & 2
baseMap.pr.12 <- createFactorMap(Fi.1, col.points = DESIGN_1$rows$PerfRating$color_observ,
                             col.labels = 'darkorchid',title = "Factor scores by Performance Rating",display.labels = FALSE,
                             display.points = TRUE,axis1 = 1,axis2 = 2)  


#Calculating the means for Dim 1 & 2
data_means_PR <- PTCA4CATA::getMeans(Fi.1, ibm.emp.data$PerformanceRating)
rownames(data_means_PR) = c("PR3","PR4")


#Mapping the means
MapGroup_pr_12    <- PTCA4CATA::createFactorMap(data_means_PR,
                            # use the constraint from the main map
                            constraints = baseMap.pr.12$constraints,
                            col.points = DESIGN_1$rows$PerfRating$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,axis1 = 1,axis2 = 2,
                            col.labels = DESIGN_1$rows$PerfRating$color_groups,
                            text.cex = 6)

#Adding all the plots
pr_1_2 <- baseMap.pr.12$zeMap + x_y_labels_1_2+MapGroup_pr_12$zeMap_dots + MapGroup_pr_12$zeMap_text

#dev.new()

print(pr_1_2)


```

#### Confidence Intervals
```{r}
# Confidence Intervals ----
# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(res_pca_d2$ExPosition.Data$fi, 
                                 design = DESIGN_1$rows$PerfRating$color_observ,
                                 niter = 100,
                                 suppressProgressBar = TRUE)


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            p.level = .95
)

PR.withCI <-  pr_1_2 +  GraphElli

#dev.new()
print(PR.withCI)

```

#### Between Component 2 & 3:
```{r}
#Plotting Factor Scores for Dim 1 & 2
baseMap.pr.23 <- createFactorMap(Fi.1, col.points = DESIGN_1$rows$PerfRating$color_observ,
                             col.labels = 'darkorchid',title = "Factor scores by Performance Rating",display.labels = FALSE,
                             display.points = TRUE,axis1 = 2,axis2 = 3)  

#Mapping the means
MapGroup_pr_23    <- PTCA4CATA::createFactorMap(data_means_PR,
                            # use the constraint from the main map
                            constraints = baseMap.pr.23$constraints,
                            col.points = DESIGN_1$rows$PerfRating$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,axis1 = 2,axis2 = 3,
                            col.labels = DESIGN_1$rows$PerfRating$color_groups,
                            text.cex = 6)

#Adding all the plots
pr_2_3 <- baseMap.pr.23$zeMap +MapGroup_pr_23$zeMap_dots + MapGroup_pr_23$zeMap_text#+x_y_labels_2_3

#dev.new()
print(pr_2_3)
```


```{r}
color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new.ibm.emp.data))
col4Levels <- data4PCCAR::coloringLevels(rownames(Fj), color4Var)
col4Labels <- col4Levels$color4Levels

varCtr <- data4PCCAR::ctr4Variables(Cj)
#rownames(color4Var) <- rownames(varCtr)
```


### Loadings
#### Loadings for Component 1 vs Component 2:

```{r}
col4Levels.imp <- data4PCCAR::coloringLevels(rownames(Fj),col4ImportantVar)

baseMap.j12 <- createFactorMap(Fj, #constraints = constraints.sym,
                             col.points = col4Levels.imp$color4Levels,col.labels = col4Levels.imp$color4Levels,
                             display.labels = TRUE,display.points = TRUE,text.cex = 3,force=2,cex=1,
                              title = "J Map for Dimension 1 & 2",axis1=1,axis2=2)
dev.new()
print(baseMap.j12$zeMap + x_y_labels_1_2)
```


#### Loadings for Component 2 vs Component 3
```{r}
baseMap.j23 <- createFactorMap(Fj, #constraints = constraints.sym,
                             col.points = col4Levels.imp$color4Levels,col.labels = col4Levels.imp$color4Levels,
                             display.labels = TRUE,display.points = TRUE,text.cex = 3,force=2,cex=1,
                              title = "J Map for Dimension 2 & 3",axis1=2,axis2=3)
dev.new()
print(baseMap.j23$zeMap + x_y_labels_1_2)
```



## Inferences

### Bootstrap Bars

We are trying to find out the significance of the different variables used in the analysis, by using bootstrap method to generate many samples from the same dataset.  

As we run the inference bootstraping takes place within and we get the bootstrap ratios for each variables.
The plot will show bootstrap ratio of each variables for each component.  

The ones greater than 2 will be shown as significant and the if the variable pops out in the top half it coressponds it is significant towards positive side of the component and otherwise.  

```{r}

BR <- res_pca.inf_d1$Inference.Data$fj.boots$tests$boot.ratios
laDim = 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                        threshold = 2,
                        font.size = 5,
                   #color4bar = gplots::col2hex(res_pca.inf_d1$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',horizontal = FALSE
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
dev.new()
print(ba001.BR1)
#
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(res_pca.inf_d1$Plotting.Data$fj.col), # we need hex code
                            main = paste0(
                              'Bootstrap ratio for Dimension:',laDim),
                ylab = 'Bootstrap ratios',horizontal = FALSE
)
#dev.new()
print(ba002.BR2)
```

For PC1 :  
- Age,MonthlyIncome,Job Level, NumofComapniesWorked,TotalWorkingYears and other year realted variables are significant in     -ve   half of component.  
- Performance Rating and Percent Salary Hike are on +ve side of the component  
For PC2 :  
- DistFrmHome,PerformanceRating,PercentSalaryHike,YearsAtCompany,YearsinCurrentRole,YearsWithCurrMangr are significant in     -ve half of component  
- StockOptionLevel, NumofCompaniesWorked,Job & Relationship Satisfaction are on +ve side of the component.  






##  Conclusions
- Factor score when plotted between PC1 and PC2,colored by Job Level showed distribution along the PC1  
- Factor score when plotted between PC2 and PC3,colored by Job Level didnt show any distirbution  

- Factor score when plotted between PC1 and PC2,colored by Performance Rating, did not show any distribution  
- Factor score when plotted between PC2 and PC3,colored by Performance Rating, did not show any distribution  

- Loading for PC1 and PC2, Percent Salary Hike & Performance Rating were inversely correlated to NUm of Companies worked &    Relationship Satisfaction . Also Year related variables are orthogonal to above 2 groups.  

- Loading for PC2 and PC3, Num of companies worked & Age were inversely correlated to Environment Satisfaction & Year related variables. Also in this case Performance rating and percent salary hike are orthogonal to above 2 groups.


