---
title: "BADA"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "11/5/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
graphics.off()


library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)
#install.packages('TExPosition')
library(TExPosition)
library(scales)

library(TInPosition)
library(officer)
library(flextable)
library(rvg)

file4PLSCfunctions <- 'inferences4PLSC.R'
# Source the function file:
source(file4PLSCfunctions)

```

# Method : BADA (Barycentric Discriminant Analysis)
Barycentric discriminant analysis (BADA) is a robust version of discriminant analysis that
is used to assign, to pre-defined groups (also called categories), observations described by
multiple variables. 
By contrast with traditional discriminant analysis, BADA can be used
even when the number of observations is smaller than the number of variables—This makes
BADA particularly suited for the analysis of Big Data.

## Data set: IBM-HR-Emplyee-WithAttrition

The dataset given is regarding IBM employees, available to the HR section of the company to take decisions regarding attrition.

It consists of data of 237 employees (rows), with 32 variables(columns) defining each employee based on position,education,department etc.


```{r data_set,echo=FALSE,include=FALSE}
# Dataset ----
ibm.emp.data <- read.xlsx("IBM-HR-Emplyee-WithAttrition.xlsx", sheetName = "Sheet1",header=TRUE)
rownames(ibm.emp.data) <- ibm.emp.data$Subj
ibm.emp.data <- ibm.emp.data[-c(1)]
cols <- c('Age','Gender','Education','MonthlyIncome','TotalWorkingYears','JobSatisfaction')
head(ibm.emp.data[cols])

```



```{r}

new.ibm.emp.data <- ibm.emp.data[sapply(ibm.emp.data, function(x) !is.factor(x))]


```

### Heat Map

After removing the factor variables we have mostly numeric ones, but in those also we have some ordinal ones,
and few have very higher range than others like income etc.  

So I tried plotting different heatmaps including ordinal ones, income ones (with log), and the remaining quant ones together.  

- Income Map : Monthly is higher than Daily rate which is again higher than Hourly Rate,Monthly rate seems to be                      higher throughout.  
- Ordinal Ones : Employees on higher Job Level have stock options
                  Employees with good work life balance have better satisfaction all through  
                 

```{r}

color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new.ibm.emp.data))

corrMatBurt.list <- phi2Mat4BurtTable(new.ibm.emp.data)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corr4MCA.r <- corrplot::corrplot(
as.matrix(corrMatBurt.list$phi2.mat),
method="color", col=col(200),
type="upper",
addCoef.col = "black", # Add coefficient of correlation
tl.col=color4Var,
tl.srt = 45, #Text label color and rotation
number.cex = .5,
diag = TRUE # needed to have the color of variables correct
)



```



```{r DESIGN,echo=FALSE,include=FALSE}
#Create an empty list, called DESIGN
JL <- ibm.emp.data$JobLevel
JLU <- unique(ibm.emp.data$JobLevel)

DESIGN <- list()
#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$JobLevel$labels <- unique(ibm.emp.data$JobLevel)
#Storing the categorical values in a vector
DESIGN$rows$JobLevel$vec <- ibm.emp.data$JobLevel
#Convert the vector to a matrix
DESIGN$rows$JobLevel$mat <- makeNominalData(as.matrix(DESIGN$rows$JobLevel$vec))

```



```{r colors_2,echo=FALSE,include=FALSE}

#As our group has 5 levels we specify same number of colors
DESIGN$rows$JobLevel$color_groups <- c("red", "green", "magenta", "blue","yellow")
#plot(c(1:5), pch=15, cex=4, col=DESIGN$rows$JobLevel$color_groups) 

#First, copy the group names
DESIGN$rows$JobLevel$color_observ <- as.matrix(DESIGN$rows$JobLevel$vec)

#Then, for each group, replace the group name with the group's color
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="1")]  <- DESIGN$rows$JobLevel$color_groups[1]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="2")]  <- DESIGN$rows$JobLevel$color_groups[2]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="3")]  <- DESIGN$rows$JobLevel$color_groups[3]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="4")]  <- DESIGN$rows$JobLevel$color_groups[4]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="5")]  <- DESIGN$rows$JobLevel$color_groups[5]

```



## RUnning BADA

```{r}

#new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(JobLevel,PerformanceRating))
new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(JobLevel))
#new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(PerformanceRating))

resBADA <- tepBADA(DATA = new.ibm.emp.data1,
                   scale = 'SS1', center = TRUE,
                   #DESIGN = DESIGN$rows$JobLevel$color_observ,
                   DESIGN = JL,
                   make_design_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL, graphs =  FALSE)

resBADA.inf <- tepBADA.inference.battery(DATA = new.ibm.emp.data1,
                   scale = 'SS1', center = TRUE,
                   #DESIGN = DESIGN$rows$JobLevel$color_observ  ,
                   DESIGN = JL,
                   make_design_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL,
                   graphs = FALSE,
                   test.iters = 100)
```


### Scree Plot
```{r}
# A look at the scree
#dev.new()
PlotScree(ev = resBADA$TExPosition.Data$eigs,
          #p.ev = resBADA.inf$Inference.Data$components$p.vals,# max.ev = NULL, alpha = 0.05,
          #col.ns = "#006D2C", col.sig = "#54278F",
          title = "Explained Variance per Dimension",plotKaiser = TRUE)
BADAScree <- recordPlot()

```

```{r,echo=FALSE}
#---------------------------------------------------------------------
# To minimize typing:
Fi <- resBADA$TExPosition.Data$fi
Fk <- resBADA$TExPosition.Data$fii
Fj <- resBADA$TExPosition.Data$fj
Cj <- resBADA$ExPosition.Data$cj
Eigs <- resBADA$TExPosition.Data$eigs
tau <- resBADA$TExPosition.Data$t

labels4BADA_12 <- createxyLabels.gen(x_axis = 1, y_axis = 2,
              lambda = Eigs,tau =tau)

#---------------------------------------------------------------------
```


### Loading plots
```{r}

#---------------------------------------------------------------------
# A graph for the J set
baseMap.j <- PTCA4CATA::createFactorMap(Fj,
                            col.points   = 'darkorchid',
                            alpha.points =  .3,
                            alpha.labels =  .7,
                            col.labels   = 'darkorchid')
# A graph for the J-set
aggMap.j <- baseMap.j$zeMap_background + # background layer
  baseMap.j$zeMap_dots + baseMap.j$zeMap_text # dots & labels
# We print this Map with the following code
#dev.new()
print(aggMap.j + labels4BADA_12)
```

```{r}
# De we want to add lines?
# PTCA4CATA does not have arrowed lines,
#    but this line of code create them
zeLines <- ggplot2::annotate("segment", x = c(0), y = c(0),
                    xend = Fj[,1],
                    yend = Fj[,2],
                    color = 'darkorchid',
                    alpha = .5,
                    arrow = arrow(length = unit(.3, "cm") ) )
# Create the map by adding background, labels, and arrows:
aggMap.j.arrows <- baseMap.j$zeMap_background +
                                      zeLines + baseMap.j$zeMap_text
#dev.new()
print(aggMap.j.arrows + labels4BADA_12)
```



## I Set map
```{r}
#---------------------------------------------------------------------
# Graphs for the I set



baseMap.i <- createFactorMap(Fk, 
                             col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Factor scores by Job Level",display.labels = FALSE)  

aggMap.i <- baseMap.i$zeMap_background + baseMap.i$zeMap_dots

print(baseMap.i$zeMap +labels4BADA_12)

#dev.new()
#print(aggMap.i)
```

## Means
```{r}

data_means <- PTCA4CATA::getMeans(Fk, ibm.emp.data$JobLevel)
rownames(data_means) = c("JL5","JL2","JL3","JL4","JL1")

# the map
MapGroup    <- PTCA4CATA::createFactorMap(data_means,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)


# The map with observations and group means
aggMap.i.withMeans <- aggMap.i+
  MapGroup$zeMap_dots + MapGroup$zeMap_text
#dev.new()
print(aggMap.i.withMeans + labels4BADA_12)
#---------------------------------------------------------------------
```


## Tolerance Interval
```{r}
#---------------------------------------------------------------------
# Create 75% Tolerance interval polygons
#
GraphTI.Hull.90 <- MakeToleranceIntervals(Fk,
                                  ibm.emp.data$JobLevel,
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = DESIGN$rows$JobLevel$color_groups,
                                  line.size = .5, line.type = 3,
                                  alpha.ellipse = .2,
                                  alpha.line = .4,
                                  p.level = .75, # 75% TI
                                  type = 'hull' #
                                          # use 'hull' for convex hull
)
#---------------------------------------------------------------------
# Create the map
aggMap.i.withHull <- aggMap.i +
  GraphTI.Hull.90 + MapGroup$zeMap_dots +
  MapGroup$zeMap_text +  MapGroup$zeMap_dots
#---------------------------------------------------------------------
# Plot it!
#dev.new()
print(aggMap.i.withHull + labels4BADA_12)
#---------------------------------------------------------------------
```

### Confusion Matrix
```{r}
fixedCM   <-   resBADA.inf$Inference.Data$loo.data$fixed.confuse
looedCM   <- resBADA.inf$Inference.Data$loo.data$loo.confuse

print(fixedCM)
print(looedCM)
```

### Confidence Intervals
```{r}
# Confidence Intervals ----


# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(Fk,
                                 design = ibm.emp.data$JobLevel,
                                 niter = 100,
                                 suppressProgressBar = TRUE)


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = DESIGN$rows$JobLevel$color_groups,
                            
                            p.level = .95
)

a004.Map.I.withCI <-  aggMap.i + MapGroup$zeMap_text +  GraphElli + labels4BADA_12

print(a004.Map.I.withCI)

```

## Inference
```{r include=FALSE,eval=FALSE}
#---------------------------------------------------------------------
# Inferences
#

#---------------------------------------------------------------------
# Confusion matrices
# To be saved as table
fixedCM   <-   resBADA.inf$Inference.Data$loo.data$fixed.confuse
looedCM   <- resBADA.inf$Inference.Data$loo.data$loo.confuse

#---------------------------------------------------------------------
# Create Confidence Interval Plots
BootCube <- resBADA.inf$Inference.Data$boot.data$fi.boot.data$boots
#dimnames(BootCube)[[2]] <- c("Dimension 1","Dimension 2")
# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- MakeCIEllipses(BootCube[,1:2,],
                  names.of.factors = c("Dimension 1","Dimension 2"),
                  #col = rownames(BootCubeBootCube[,c(1,2),1]),
                  p.level = .95
)
#---------------------------------------------------------------------
# create the I-map with Observations, means and confidence intervals
#
aggMap.i.withCI <-  aggMap.i +  GraphElli + MapGroup$zeMap_text + labels4BADA_12
#---------------------------------------------------------------------
# plot it!
#dev.new()
print(aggMap.i.withCI)
```




## Contribution Plot

```{r}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrJ <- resBADA$TExPosition.Data$cj * sign(resBADA$TExPosition.Data$fj)

b003.ctrJ.bada.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(resBADA$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b003.ctrJ.bada.1)


b004.ctrJ.bada.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                           threshold = 1 / NROW(signed.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(resBADA$Plotting.Data$fj.col), # we need hex code
                           main = 'Variable Contributions (Signed): Dimension 2',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b004.ctrJ.bada.2)

```

**Age, Monthly Income,Total Working Years and other Year realted variables lie on the -ve side of dimension 1**
**Performance Rating, Percent Salary Hike, Train Times Last Year on +ve side of dimension 1**
**DailyRate, Monthly Rate,WorkLifeBalance, EnvironmentStifaction on -ve of DImension 2**
**Hourly Rate, DIstFrm Home, Education, Job Invovlement Job Satisfactio on +ve of Dimesnion 2**

## Bootstrap Bars

```{r}

BADA.BR <- resBADA.inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios
laDim = 1
ba001.bada.BR1 <- PrettyBarPlot2(BADA.BR[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resBADA$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios'
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.bada.BR1)
#
laDim = 2
ba002.bada.BR2 <- PrettyBarPlot2(BADA.BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            color4bar = gplots::col2hex(resBADA$Plotting.Data$fj.col), # we need hex code
                            main = paste0(
                              'Bootstrap ratio for Dimension:',laDim),
                ylab = 'Bootstrap ratios'
)
#dev.new()
print(ba002.bada.BR2)
```
**Age, Monthly Income,Total Working Years and other Year realted variables lie on the -ve side of dimension 1**  
**Performance Rating, Percent Salary Hike, Train Times Last Year on +ve side of dimension 1**  
**Monthly Rate,WorkLifeBalance,Perf Rating on -ve of DImension 2**  



## Conclusion
- BADA was applied and categorized based on Job Level (which) had 5 levels.  
- Factor scores were divided amongst the Job levels, with less overlapping.  
- Based on which we can predict the levels of supplemental observations.
