
---
title: "CA on Bird_Brain"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "10/2/2018"
output:
  word_document: default
  html_document: default
---

```{r CA_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
graphics.off()

library(ExPosition)
library(InPosition)
# use the last version of PTCA
#devtools::install_github('HerveAbdi/PTCA4CATA')
library(PTCA4CATA)
library(corrplot)
library(ggplot2)
#install.packages('gplots')
# also install data4PCCAR
#devtools::install_github('HerveAbdi/data4PCCAR')
library(data4PCCAR)
```


# Correspondence Analysis
CA is a generalized version of PCA for analyzing data table with qualitative variables.
In case of PCA we divided the table into factor scores(rows) and loadings(variables), where as in CA, we derive 2 set of factor scores one for the rows and the other for the columns.
-Rows have masses representing each rows importance wrt the total table
-similarly Columns have weights each columns importance in describing each of the rows



## Dataset: Bird Brain
The Bird Brain data set  contains data about about 28 different birds species, and they are measured using 11 variables, one being categorical and remaining 10 quantitative.
One categorical variable groups these birds into different levels.
The quant variables describe the count of neurons in various structure of brain.
```{r}
#Fetching the Bird brian dataset

bb_df <- read.csv("BirdBrain.csv")

cols <- c('X','Order','Whole.Brain','Pallium','Diencephalon','Cerebellum')
head(bb_df[cols])

```



#### Working on dataset

```{r}
row.names(bb_df) <- bb_df$X
new_bb_df <- bb_df[-c(1,2)]
B <- as.matrix(new_bb_df)

```




### Heatmap 

**Heat map shows that the cerebellum part is the most crowdedwith neurons for all the species observed here.**

-Goldcrest has more density in 3 more portions of brain.

-Parrot has a rare higher density at subpallium

-While EMu has pretty musch lesser neurons than rest in all regions except cerebellum.

```{r}
#heatmap(B,Rowv=NA,Colv=NA,main="Heat map for Bird Brain",xlab="Log(num of Neuron) as per brain struture",ylab="Bird Species",margins=c(10,8) )

col4J.Bird <- prettyGraphsColorSelection(NCOL(bb_df))
c000.heatMapIJ.IBM <- makeggHeatMap4CT(new_bb_df,
                                       colorAttributes = col4J.Bird,
fontSize.x = 15)
print(c000.heatMapIJ.IBM)

```



### 1st Design Variable : Order

Selecting order as the grouping variable

```{r DESIGN_CA}
#Create an empty list, called DESIGN
DESIGN <- list()
#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$Order$labels <- unique(bb_df$Order)
#Storing the categorical values in a vector
DESIGN$rows$Order$vec <- bb_df$Order
#Convert the vector to a matrix
DESIGN$rows$Order$mat <- makeNominalData(as.matrix(DESIGN$rows$Order$vec))

```

### Assiging color to the groups
```{r DESIGN_CA_colors_2}
DESIGN$rows$Order$color_groups <- c("red", "green", "magenta", "blue","yellow","brown","orange")
#First, copy the group names
DESIGN$rows$Order$color_observ <- as.matrix(DESIGN$rows$Order$vec)
count = 0
for (val in unique(bb_df$Order)) {
  count = count + 1
  DESIGN$rows$Order$color_observ[which(DESIGN$rows$Order$vec==val)]  <- DESIGN$rows$Order$color_groups[count]
}

```




## Running CA

Running the symmetric and asymmetric as well as the inference pca
```{r}
# Running the Symmetric CA
resCA.sym  <- epCA(B, symmetric = TRUE,DESIGN = bb_df$Order)

# to run a plain CA but asymetric
resCA.asym <- epCA(B, symmetric = FALSE,DESIGN = bb_df$Order)


resCA.inf <- InPosition::epCA.inference.battery(DATA = B,
                   #scale = 'SS1', # Make sure to use 'SS1' rather than TRUE
                   DESIGN = bb_df$Order,
                   graphs =  FALSE # TRUE first pass only
                    )


# HA as sup ----
#HA.sup <- supplementaryRows(SUP.DATA = HApunct, res = resCA.sym)
# Other punct ----
#punct.sup <- supplementaryCols(SUP.DATA = otherPunct, res = resCA.sym)

```


### Scree Plot

**From the permutation test on the scree plot, we can see that 2 components are of significance**
```{r}
EigenValues <- resCA.asym$ExPosition.Data$eigs


PlotScree(ev = EigenValues, 
       p.ev =  resCA.inf$Inference.Data$components$p.vals,
       title = "Scree Plot with Permutation Test",
       plotKaiser = TRUE
          )
a001a.screePlot <- recordPlot()
#dev.new()
print(a001a.screePlot)

```


```{r,include=FALSE}
# CA graphs ----
# to make life easier ----
#Fi.a <- resCA.asym$ExPosition.Data$fi ##Additional
Fj.a <- resCA.asym$ExPosition.Data$fj
Fi   <- resCA.sym$ExPosition.Data$fi
Fj   <- resCA.sym$ExPosition.Data$fj

# constraints -----
# first get the constraints correct
constraints.sym <- minmaxHelper(mat1 = Fi, mat2  = Fj)
constraints.asym <- minmaxHelper(mat1 = Fi, mat2  = Fj.a)    ##Can add Fi.a
#constraints.sup <- minmaxHelper(mat1 = rbind(Fi, HA.sup$fii), mat2  = rbind(Fj, punct.sup$fjj) )
```

### Creating Base maps
```{r}
# Get some colors ----
#length(unique(bb_df$Order))
color4Authors <-prettyGraphsColorSelection(n.colors = nrow(Fi))
# baseMaps ----
baseMap.i <- createFactorMap(Fi, constraints = constraints.sym,
                             col.points = color4Authors,
                             col.labels = color4Authors,title = "Neuron Distribution in Birds")
#dev.new()
print(baseMap.i$zeMap)


baseMap.j <- createFactorMap(Fj, constraints = constraints.sym,
                             color.points = 'darkorchid4')
#dev.new()
print(baseMap.j$zeMap)
print(baseMap.i$zeMap + baseMap.j$zeMap_dots + baseMap.j$zeMap_text)
```



### Plotting observation for symmetric and asymmetric
```{r}


symMap  <- createFactorMapIJ(Fi,Fj,
                             col.points.i = DESIGN$rows$Order$color_observ,
                             col.labels.i = DESIGN$rows$Order$color_observ)

asymMap  <- createFactorMapIJ(Fi,Fj.a,
                              col.points.i = DESIGN$rows$Order$color_observ,
                              col.labels.i = DESIGN$rows$Order$color_observ)

labels4CA <- createxyLabels(resCA = resCA.sym)
```

## Inferences

### Contribution Plot

**Tectum,Diencephalon and Brainstem have significant contributions to the Dimension 1**
**Whereas, Subpallium,Cerebellum and Rest of Brain have more significant contributions in Dimension 2**
```{r}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrJ <- resCA.sym$ExPosition.Data$cj * sign(resCA.sym$ExPosition.Data$fj)

b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(resCA.sym$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b003.ctrJ.s.1)


b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                           threshold = 1 / NROW(signed.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(resCA.sym$Plotting.Data$fj.col), # we need hex code
                           main = 'Variable Contributions (Signed): Dimension 2',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b004.ctrJ.s.2)

```

### COntribution Plots Asymetric
```{r}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrJ.a <- resCA.asym$ExPosition.Data$cj * sign(resCA.asym$ExPosition.Data$fj)

b003.ctrJ.a.1 <- PrettyBarPlot2(signed.ctrJ.a[,1],
                         threshold = 1 / NROW(signed.ctrJ.a),
                         font.size = 5,
                         color4bar = gplots::col2hex(resCA.asym$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ.a), 1.2*max(signed.ctrJ.a))
)
#dev.new()
print(b003.ctrJ.a.1)


b004.ctrJ.a.2 <- PrettyBarPlot2(signed.ctrJ.a[,2],
                           threshold = 1 / NROW(signed.ctrJ.a),
                           font.size = 5,
                           color4bar = gplots::col2hex(resCA.asym$Plotting.Data$fj.col), # we need hex code
                           main = 'Variable Contributions (Signed): Dimension 2',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ.a), 1.2*max(signed.ctrJ.a))
)
#dev.new()
print(b004.ctrJ.a.2)

```




### Bootstrap Bars
**For Dimension 1 its shows, all the variables have significant contribution**
**FOr Dimension 2 its shows, Subpallium,Tectum,Diencephalon and Brainstem have significant contribution**

```{r}

BR <- resCA.inf$Inference.Data$fj.boots$tests$boot.ratios
laDim = 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resCA.sym$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios'
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR1)
#
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            color4bar = gplots::col2hex(resCA.sym$Plotting.Data$fj.col), # we need hex code
                            main = paste0(
                              'Bootstrap ratio for Dimension:',laDim),
                ylab = 'Bootstrap ratios'
)
#dev.new()
print(ba002.BR2)
```


### Symmetric and Asymmetric Plots

**->From the symmetric plot:**
  **- we can interpret that the Passeriformes(GOldcrest, STraling etc) have more neurons towards Brainstem & Diencephalon           compared to others.**
  **- Psittaciformes(Parrot, Parakeet) have more neuron towards Subpallium region**
  **- Corvid(Raven, MAgpie) are towards Pallium**
  **- WHile Emu, Pigeon, Owl etc ae far from others but still closer to the cerebellum region**


```{r}
# draw the maps ----


map.IJ.sym <- symMap$baseMap + symMap$I_labels + symMap$I_points +
  symMap$J_labels + symMap$J_points + labels4CA

#dev.new()
print(map.IJ.sym)


map.IJ.asym <- asymMap$baseMap + asymMap$I_labels + 
  asymMap$I_points + asymMap$J_labels + 
  asymMap$J_points + labels4CA


#dev.new()
print(map.IJ.asym)
```



```{r , include=FALSE}
# create an asymetric map with a supplementary row
#map.I.sup.asym <- asymMap$baseMap + asymMap$I_labels + 
#  asymMap$I_points +
#  asymMap$J_labels + asymMap$J_points + 
  #mapSup$I_labels + mapSup$I_points +
#  labels4CA
#dev.new()
#print(map.I.sup.asym)

```

```{r,include=FALSE}
# Create a symmetric map with sup and correct constraints
#map.IJ.sup.sym <- mapSup$baseMap + 
 # symMap$I_labels + symMap$I_points +
 # symMap$J_labels + symMap$J_points + 
  #mapSup$I_labels + mapSup$I_points + 
  #mapSup$J_labels + mapSup$J_points +
 # ggtitle('Symmetric Map with Supplementary Elements') + 
 # labels4CA
#print(map.IJ.sup.sym)



```

## Conclusion
### Component 1:  
**Rows: Passeriformes VS Psittaciformes & Corvid**
**Columns : Brainstem,Diencephalon and Tectum VS Others**
### Component 2:  
**Rows: Struthioniformes(Emu) & Columbiformes(Pigeon) & Galliformes(Junglefowl)  VS Rest**
**Columns : Rest of Brain & Subpallium VS Cerebellum & Whole Brain**

**Interpretation : Majority birds are concentrated near the whole brain, palium,telencephalon region. Few birds of Passeriformes order have neurons in Diencephalon & Brainstem. While EMu shows more neurons near the Cerebellum area.**

