
---
title: "DiCA"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "11/12/2018"
output:
  word_document: default
  html_document: default
---


```{r DiCA_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
graphics.off()


library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)
#install.packages('TExPosition')
library(TExPosition)
library(scales)

library(TInPosition)
library(officer)
library(flextable)
library(rvg)

file4PLSCfunctions <- 'inferences4PLSC.R'
# Source the function file:
source(file4PLSCfunctions)

```


# Discriminant Correspondence Analysis
DICA is an exension on CA and DA(grouping to predefined categories).
-The main idea behind DCA is to represent each group by the
sum of its observations and to perform a simple CA on the groups
by variables matrix.

## Data set: IBM-HR-Emplyee-WithAttrition

The dataset given is regarding IBM employees, available to the HR section of the company to take decisions regarding attrition.

It consists of data of 237 employees (rows), with 32 variables(columns) defining each employee based on position,education,department etc.


```{r DiCA_data_set}
# Dataset ----
ibm.emp.data <- read.xlsx("IBM-HR-Emplyee-WithAttrition.xlsx", sheetName = "Sheet1",header=TRUE)
rownames(ibm.emp.data) <- ibm.emp.data$Subj
ibm.emp.data <- ibm.emp.data[-c(1)]
cols <- c('Age','Gender','Education','MonthlyIncome','TotalWorkingYears','JobSatisfaction')
head(ibm.emp.data[cols])

```



```{r}

new.ibm.emp.data <- ibm.emp.data[sapply(ibm.emp.data, function(x) !is.factor(x))]

```



```{r DiCA_DESIGN,echo=FALSE}
#Create an empty list, called DESIGN
JL <- ibm.emp.data$JobLevel
JLU <- unique(ibm.emp.data$JobLevel)

DESIGN <- list()
#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$JobLevel$labels <- unique(ibm.emp.data$JobLevel)
#Storing the categorical values in a vector
DESIGN$rows$JobLevel$vec <- ibm.emp.data$JobLevel
#Convert the vector to a matrix
DESIGN$rows$JobLevel$mat <- makeNominalData(as.matrix(DESIGN$rows$JobLevel$vec))

```



```{r DiCA_colors_2,echo=FALSE}

#As our group has 5 levels we specify same number of colors
DESIGN$rows$JobLevel$color_groups <- c("red", "green", "magenta", "blue","yellow")
#plot(c(1:5), pch=15, cex=4, col=DESIGN$rows$JobLevel$color_groups) 

#First, copy the group names
DESIGN$rows$JobLevel$color_observ <- as.matrix(DESIGN$rows$JobLevel$vec)

#Then, for each group, replace the group name with the group's color
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="1")]  <- DESIGN$rows$JobLevel$color_groups[1]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="2")]  <- DESIGN$rows$JobLevel$color_groups[2]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="3")]  <- DESIGN$rows$JobLevel$color_groups[3]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="4")]  <- DESIGN$rows$JobLevel$color_groups[4]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="5")]  <- DESIGN$rows$JobLevel$color_groups[5]

```



### Binnig the variables
```{r include=FALSE,echo=FALSE}
new_df =data.frame()

#Age ---- NOrmal Distribution

hist(new.ibm.emp.data[,1],breaks = 25, main="Age", xlab = "Values")
#abline(v = c(28,32,39), col = "red", lwd =2) # plot the cutoff line

qts <- quantile(new.ibm.emp.data[,1])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

age <- cut(new.ibm.emp.data[,1],breaks=c(min(new.ibm.emp.data[,1])-1,qts,max(new.ibm.emp.data[,1])+1),labels=c(1,2,3,4))

cor(as.numeric(age),new.ibm.emp.data[,1],method = "spearman") #Checking for consistency


#MOnthly Income ---- 

#hist(new.ibm.emp.data[,2],breaks = 25, main="Length", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,2])[2:4]
#abline(v = qts, col = "red", lwd =2) # plot the cutoff line

monthly_inc <- cut(new.ibm.emp.data[,2],breaks=c(min(new.ibm.emp.data[,2])-1,qts,max(new.ibm.emp.data[,2])+1),labels=c(1,2,3,4))

cor(as.numeric(monthly_inc),new.ibm.emp.data[,2],method = "spearman") #Checking for consistency



#Daily Rate ---- 

hist(new.ibm.emp.data[,3],breaks = 25, main="Daily Rate", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,3])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

daily_rt <- cut(new.ibm.emp.data[,3],breaks=c(min(new.ibm.emp.data[,3])-1,qts,max(new.ibm.emp.data[,3])+1),labels=c(1,2,3,4))

cor(as.numeric(daily_rt),new.ibm.emp.data[,3],method = "spearman") #Checking for consistency


#Hourly Rate ---- 

hist(new.ibm.emp.data[,4],breaks = 25, main="Length", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,4])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

hourly_rt <- cut(new.ibm.emp.data[,4],breaks=c(min(new.ibm.emp.data[,4])-1,qts,max(new.ibm.emp.data[,4])+1),labels=c(1,2,3,4))

cor(as.numeric(hourly_rt),new.ibm.emp.data[,4],method = "spearman") #Checking for consistency



#Monthly Rate ---- 

hist(new.ibm.emp.data[,5],breaks = 25, main="Monthly Rate", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,5])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

monthly_rt <- cut(new.ibm.emp.data[,5],breaks=c(min(new.ibm.emp.data[,5])-1,qts,max(new.ibm.emp.data[,5])+1),labels=c(1,2,3,4))

cor(as.numeric(monthly_rt),new.ibm.emp.data[,5],method = "spearman") #Checking for consistency



#Distance From HOme ---- 


hist(new.ibm.emp.data[,6],breaks = 25, main="Distance From Home", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,6])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

dist_frm_Hm <- cut(new.ibm.emp.data[,6],breaks=c(min(new.ibm.emp.data[,6])-1,qts,max(new.ibm.emp.data[,6])+1),labels=c(1,2,3,4))

cor(as.numeric(dist_frm_Hm),new.ibm.emp.data[,6],method = "spearman") #Checking for consistency



#Num of companies worked----

#unique(new.ibm.emp.data[,12])
#table(new.ibm.emp.data[,12])
hist(new.ibm.emp.data[,12],breaks = 25, main="Num of companies worked", xlab = "Values")
abline(v = c(0,1,5), col = "red", lwd =2) # plot the cutoff line

num_comp <- cut(new.ibm.emp.data[,12],breaks=c(min(new.ibm.emp.data[,12])-1,c(0,1,5),max(new.ibm.emp.data[,12])+1),labels=c(1,2,3,4))

cor(as.numeric(num_comp),new.ibm.emp.data[,12],method = "spearman") #Checking for consistency


#Percent salary hike----

#unique(new.ibm.emp.data[,13])
#table(new.ibm.emp.data[,13])

hist(new.ibm.emp.data[,13],breaks = 25, main="Percent Salary Hike", xlab = "Values")
abline(v = c(11,13,17,20), col = "red", lwd =2) # plot the cutoff line

per_hike <- cut(new.ibm.emp.data[,13],breaks=c(min(new.ibm.emp.data[,13])-1,c(11,13,17,20),max(new.ibm.emp.data[,13])+1),labels=c(1,2,3,4,5))

cor(as.numeric(per_hike),new.ibm.emp.data[,13],method = "spearman") #Checking for consistency



#Total Working Years----

#unique(new.ibm.emp.data[,14])
#table(new.ibm.emp.data[,14])

hist(new.ibm.emp.data[,14],breaks = 25, main="Total Working Years", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,14])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

tot_Wrk_yrs <- cut(new.ibm.emp.data[,14],breaks=c(min(new.ibm.emp.data[,14])-1,qts,max(new.ibm.emp.data[,14])+1),labels=c(1,2,3,4))

cor(as.numeric(tot_Wrk_yrs),new.ibm.emp.data[,14],method = "spearman") #Checking for consistency


#Train Time Last Year----
#unique(new.ibm.emp.data[,15])
#table(new.ibm.emp.data[,15])
hist(new.ibm.emp.data[,15],breaks = 25, main="Train Time since Last Year", xlab = "Values")

abline(v = c(1,2,4), col = "red", lwd =2) # plot the cutoff line

train_tim_yrs <- cut(new.ibm.emp.data[,15],breaks=c(min(new.ibm.emp.data[,15])-1,c(1,2,4),max(new.ibm.emp.data[,15])+1),labels=c(1,2,3,4))

cor(as.numeric(train_tim_yrs),new.ibm.emp.data[,15],method = "spearman") #Checking for consistency


#Years at Company----

#unique(new.ibm.emp.data[,17])
#table(new.ibm.emp.data[,17])
#hist(new.ibm.emp.data[,17],breaks = 25, main="Length", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,17])[2:4]
#abline(v = c(1,2,4), col = "red", lwd =2) # plot the cutoff line

yr_at_comp <- cut(new.ibm.emp.data[,17],breaks=c(min(new.ibm.emp.data[,17])-1,qts,max(new.ibm.emp.data[,17])+1),labels=c(1,2,3,4))

cor(as.numeric(yr_at_comp),new.ibm.emp.data[,17],method = "spearman") #Checking for consistency


#Years In Current Role----

#unique(new.ibm.emp.data[,18])
#table(new.ibm.emp.data[,18])
hist(new.ibm.emp.data[,18],breaks = 25, main="Years In Current Role", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,18])[2:4]
abline(v = qts, col = "red", lwd =2) # plot the cutoff line

yr_in_role <- cut(new.ibm.emp.data[,18],breaks=c(min(new.ibm.emp.data[,18])-1,qts,max(new.ibm.emp.data[,18])+1),labels=c(1,2,3,4))

cor(as.numeric(yr_in_role),new.ibm.emp.data[,18],method = "spearman") #Checking for consistency




#Years since last promotion----

#unique(new.ibm.emp.data[,19])
#table(new.ibm.emp.data[,19])
#hist(new.ibm.emp.data[,19],breaks = 25, main="Length", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,19])[2:4]
#abline(v = c(1,2,4), col = "red", lwd =2) # plot the cutoff line

yr_since_Promo <- cut(new.ibm.emp.data[,19],breaks=c(min(new.ibm.emp.data[,19])-1,qts,max(new.ibm.emp.data[,19])+1),labels=c(1,2,3,4))

cor(as.numeric(yr_since_Promo),new.ibm.emp.data[,19],method = "spearman") #Checking for consistency


#Years with Manager----

#unique(new.ibm.emp.data[,20])
#table(new.ibm.emp.data[,20])
#hist(new.ibm.emp.data[,20],breaks = 25, main="Length", xlab = "Values")
qts <- quantile(new.ibm.emp.data[,20])[2:4]
#abline(v = c(1,2,4), col = "red", lwd =2) # plot the cutoff line

yr_wid_mgr <- cut(new.ibm.emp.data[,20],breaks=c(min(new.ibm.emp.data[,20])-1,qts,max(new.ibm.emp.data[,20])+1),labels=c(1,2,3,4))

cor(as.numeric(yr_wid_mgr),new.ibm.emp.data[,20],method = "spearman") #Checking for consistency


#new_df = cbind(age_recode,monthly_recode,daily_recode,hourly_recode,monthly_rate_recode,dist_frm_Hm,num_comp,per_hike,tot_Wrk_yrs,train_tim_yrs,yr_at_comp,yr_in_role,yr_in_role,yr_wid_mgr)


##ORDINAL ------- *********************************************************

#Performance Rating ---- 

#unique(new.ibm.emp.data[,7])
#hist(new.ibm.emp.data[,7],breaks = 25, main="Length", xlab = "Values")
perf_rat = new.ibm.emp.data[,7]


#Education ---- 
#unique(new.ibm.emp.data[,8])
education = new.ibm.emp.data[,8]


#Job Involvement ---- 

#unique(new.ibm.emp.data[,9])
job_involve = new.ibm.emp.data[,9]

#Job Level----
#unique(new.ibm.emp.data[,10])
#job_level = new.ibm.emp.data[,10]


#Stock Level Level----
#unique(new.ibm.emp.data[,11])

stock_level = new.ibm.emp.data[,11]

#Work Life Balance----
#unique(new.ibm.emp.data[,16])
#table(new.ibm.emp.data[,16])
wrk_lif_bal = new.ibm.emp.data[,16]

#Env Satisfaction----
#unique(new.ibm.emp.data[,21])
#table(new.ibm.emp.data[,21])
env_sat = new.ibm.emp.data[,21]

#Job Satisfaction  22
job_sat = new.ibm.emp.data[,22]

#Relation Satisfaction  23
rel_sat = new.ibm.emp.data[,23]

new_df = as.data.frame(cbind(age,monthly_inc,daily_rt,hourly_rt,monthly_rt,dist_frm_Hm,num_comp,per_hike,tot_Wrk_yrs,train_tim_yrs,yr_at_comp,yr_in_role,yr_since_Promo,yr_wid_mgr,education,perf_rat,job_involve,stock_level,wrk_lif_bal,env_sat,job_sat,rel_sat))
#job_level
#

dim(new_df)
```


## Running Dica
```{r}

resDICA <- tepDICA(DATA = new_df,
                   DESIGN = JL,
                   make_design_nominal = TRUE,
                   make_data_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL, graphs =  FALSE)

resDICA.inf <- tepDICA.inference.battery(DATA = new_df,
                   DESIGN = JL,
                   make_design_nominal = TRUE,
                   make_data_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL,
                   graphs = FALSE,
                   test.iters = 100,
                   critical.value = 2)
```

### DiCA Scree Plot
```{r}
# A look at the scree
#dev.new()
PlotScree(ev = resDICA$TExPosition.Data$eigs,
          p.ev = resDICA.inf$Inference.Data$components$p.vals, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F",
          title = "Explained Variance per Dimension - DiCA",plotKaiser = TRUE)
DICAScree <- recordPlot()

```

### Heat Map
```{r echo=FALSE}

color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new_df))

corrMatBurt.list <- phi2Mat4BurtTable(new_df)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corr4MCA.r <- corrplot::corrplot(
as.matrix(corrMatBurt.list$phi2.mat),
method="color", col=col(200),
type="upper",
addCoef.col = "black", # Add coefficient of correlation
tl.col=color4Var,
tl.srt = 45, #Text label color and rotation
number.cex = .5,
diag = TRUE # needed to have the color of variables correct
)



```

```{r echo=FALSE}
#---------------------------------------------------------------------
# To minimize typing:
Fi <- resDICA$TExPosition.Data$fi
Fj <- resDICA$TExPosition.Data$fj
Fk <- resDICA$TExPosition.Data$fii
Eigs <- resDICA$TExPosition.Data$eigs
tau <- resDICA$TExPosition.Data$t
Cj <- resDICA$TExPosition.Data$cj


color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(new_df))
col4Levels <- data4PCCAR::coloringLevels(rownames(Fj), color4Var)
col4Labels <- col4Levels$color4Levels

varCtr <- data4PCCAR::ctr4Variables(Cj)



absCtrVar <- as.matrix(varCtr) %*% diag(Eigs)
varCtr12 <- (absCtrVar[,1] + absCtrVar[,2]) /(Eigs[1] + Eigs[2])
importantVar <- (varCtr12 >= 1 / length(varCtr12))
col4ImportantVar <- color4Var
col4NS <- 'white'
col4ImportantVar[!importantVar] <- col4NS
#Color for importance in  variables
col4Levels.imp <- data4PCCAR::coloringLevels(rownames(Fj),col4ImportantVar)

#Labels
labels4DICA_12 <- createxyLabels.gen(x_axis = 1, y_axis = 2,
              lambda = Eigs,tau =tau)

labels4DICA_23 <- createxyLabels.gen(x_axis = 2, y_axis = 3,
              lambda = Eigs,tau =tau)

dica.constraints<- minmaxHelper(mat1 = Fi, mat2  = Fj)
#---------------------------------------------------------------------
```

### Factor Map

```{r}


#---------------------------------------------------------------------
#---------------------------------------------------------------------
# A graph for the J set
baseMap.dica.j <- PTCA4CATA::createFactorMap(Fj,
                            col.points = col4Levels.imp$color4Levels,col.labels = col4Levels.imp$color4Levels,
                            alpha.points =  .3,display.labels = TRUE,display.points = TRUE,text.cex = 2.5,force=2,cex=1,
                              title = "J Map for Dimension 1 & 2"
                            )

lines4J <- addLines4MCA(Fj, col4Var = col4Levels.imp$color4Variables, size = 1)

# A graph for the J-set
aggMap.j <- baseMap.dica.j$zeMap_background + # background layer
            baseMap.dica.j$zeMap_dots + baseMap.dica.j$zeMap_text + lines4J + labels4DICA_12



print(aggMap.j)
```

### Adding Lines
```{r eval =FALSE,include=FALSE}
# De we want to add lines?
# PTCA4CATA does not have arrowed lines,
#    but this line of code create them
zeLines.dica <- ggplot2::annotate("segment", x = c(0), y = c(0),
                    xend = Fj[,1],
                    yend = Fj[,2],
                    color = 'darkorchid',
                    alpha = .5,
                    arrow = arrow(length = unit(.3, "cm") ) )
# Create the map by adding background, labels, and arrows:
aggMap.j.arrows <- baseMap.dica.j$zeMap_background +
                                      zeLines.dica + baseMap.dica.j$zeMap_text
#dev.new()
print(aggMap.j.arrows)
```

### I Set map
```{r}
col4DICA <- DESIGN$rows$JobLevel$color_observ

#---------------------------------------------------------------------
# Graphs for the I set
#---------------------------------------------------------------------
# Create the base map
baseMap.dica.i <- PTCA4CATA::createFactorMap(Fk,
                                        col.points   = DESIGN$rows$JobLevel$color_observ,
                                        alpha.points =  .8, title = "I Map (Job Level)- DiCA")
#---------------------------------------------------------------------
# We are not interested about the labels here
#  so we will use only the dots
#  in all I-set graph.
#---------------------------------------------------------------------
# Plain map with color for the I-set
aggMap.jl_12 <- baseMap.dica.i$zeMap_background + baseMap.dica.i$zeMap_dots + labels4DICA_12
#---------------------------------------------------------------------
# print this Map
#dev.new()
print(aggMap.jl_12)
```
**Here we are getting a shape similar to Horsehsoe (Goodmans shape)**  

## Inferences
### Means
```{r}
#---------------------------------------------------------------------
# a vector of color for the means
col4Means <- unique(col4DICA)


data_means <- PTCA4CATA::getMeans(Fk, ibm.emp.data$JobLevel)
rownames(data_means) = c("JL5","JL2","JL3","JL4","JL1")

# the map
MapGroup    <- PTCA4CATA::createFactorMap(data_means,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)


# The map with observations and group means
aggMap.i.withMeans <- aggMap.jl_12+
  MapGroup$zeMap_dots + MapGroup$zeMap_text
dev.new()
print(aggMap.i.withMeans)
#---------------------------------------------------------------------
```


### Tolerance Interval
```{r}
#---------------------------------------------------------------------
# Create 75% Tolerance interval polygons
#
GraphTI.Hull.90 <- MakeToleranceIntervals(Fk,
                                 ibm.emp.data$JobLevel,
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = DESIGN$rows$JobLevel$color_groups,
                                  line.size = .5, line.type = 3,
                                  alpha.ellipse = .2,
                                  alpha.line = .4,
                                  p.level = .75, # 75% TI
                                  type = 'hull' #
                                          # use 'hull' for convex hull
)
#---------------------------------------------------------------------
# Create the map
aggMap.i.withHull <- aggMap.jl_12 +
  GraphTI.Hull.90 + MapGroup$zeMap_dots +
  MapGroup$zeMap_text +  MapGroup$zeMap_dots
#---------------------------------------------------------------------
# Plot it!
#dev.new()
print(aggMap.i.withHull)
#---------------------------------------------------------------------
```


```{r}
#---------------------------------------------------------------------
# Inferences
#

#---------------------------------------------------------------------
# Confusion matrices
# To be saved as table
fixedCM   <-   resDICA.inf$Inference.Data$loo.data$fixed.confuse
looedCM   <- resDICA.inf$Inference.Data$loo.data$loo.confuse

# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(Fk,
                                 design = ibm.emp.data$JobLevel,
                                 niter = 100,
                                 suppressProgressBar = TRUE)

#_____________________________________________________________________
# Create the ellipses
#_____________________________________________________________________
# Create Confidence Interval Plots
# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = DESIGN$rows$JobLevel$color_groups,
                            p.level = .95
)


#---------------------------------------------------------------------
# create the I-map with Observations, means and confidence intervals
#
aggMap.i.withCI <-  aggMap.jl_12 +  GraphElli + MapGroup$zeMap_text
#---------------------------------------------------------------------
# plot it!
#dev.new()
print(aggMap.i.withCI)
```

### Contribution Plot

```{r}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrJ <- resDICA$TExPosition.Data$cj * sign(resDICA$TExPosition.Data$fj)

b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(resDICA$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b003.ctrJ.s.1)


b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                           threshold = 1 / NROW(signed.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(resDICA$Plotting.Data$fj.col), # we need hex code
                           main = 'Variable Contributions (Signed): Dimension 2',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b004.ctrJ.s.2)

```

**## For Dimension 1:**
**-Age3,MonthlyInc3,Tot_Wrk_Yr_(2,3),numofComp2,yr_at_comp3,Job_sat3 all these lie on the +ve half of the dimension 1.**
**-MonthlyInc(1,2),distFrmHm_4,numofComp3,per_hike2 all these lie on the -ve half of the dimension 1.**

**## For Dimension 2:**
**Also shows a similar trend for almost all the same variables**




```{r eval=FALSE,echo=FALSE}

BR <- resDICA.inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios

laDim = 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                        threshold = 2,
                        font.size = 1,
                   #color4bar = gplots::col2hex(resDICA.inf$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',
                  ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR1)
#
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(resDICA.inf$Plotting.Data$fj.col), # we need hex code
                            main = paste0(
                              'Bootstrap ratio for Dimension:',laDim),
                ylab = 'Bootstrap ratios'
)
#dev.new()
print(ba002.BR2)
```

## Conclusion
- Dica also shows a similar pattern as for MCA when colored by Job Level

