
---
title: "MFA"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "11/19/2018"
output:
  html_document: default
  word_document: default
---

```{r MFA_setup, include=FALSE}
rm(list = ls())
graphics.off()

knitr::opts_chunk$set(echo = TRUE)

#install.packages('DistatisR')

library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)

library(scales)

library(TInPosition)
library(officer)
library(flextable)
library(rvg)

library(useful)

library(MExPosition)
```


# Multiple Factor Analysis
MFA is used to handle multiple tables, that have different variables measuring the same observations, or same set of variables on different set of observations.
- First, each table is normalized by 1st singular value (obtained by doing PCA on the same table)
- All these normalized tables are merged together and then again a PCA is done on this, which gives us factor score and loading on these variables.


## Data set: IBM-HR-Emplyee-WithAttrition

The dataset given is regarding IBM employees, available to the HR section of the company to take decisions regarding attrition.

It consists of data of 237 employees (rows), with 32 variables(columns) defining each employee based on position,education,department etc.


```{r MFA_data_set}
# Dataset ----
ibm.emp.data <- read.xlsx("IBM-HR-Emplyee-WithAttrition.xlsx", sheetName = "Sheet1",header=TRUE)
rownames(ibm.emp.data) <- ibm.emp.data$Subj
ibm.emp.data <- ibm.emp.data[-c(1)]
cols <- c('Age','Gender','Education','MonthlyIncome','TotalWorkingYears','JobSatisfaction')
head(ibm.emp.data[cols])

```
### Removing Factor variables
```{r}
new.ibm.emp.data <- ibm.emp.data[sapply(ibm.emp.data, function(x) !is.factor(x))]
```


### Splitting the data into 3 tables, 2 for the correaltion operation, 3rd for the deisgn
```{r}
new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(JobLevel))

dim(new.ibm.emp.data1)
colnames(new.ibm.emp.data1)
#a <- c('1','1','1','1','1','1','2','2','2','2','2','2','3','3','3','3','3','3','3','3','3','3')
#column.design <- c(1,1,1,1,1,1,2,2,2,2,2,2,3,3,3,3,3,3,3,3,3,3)
column.design <- c(1,1,1,1,1,1,2,2,2,2,2,2,3,3,2,3,3,3,3,2,2,2)

column.design = t(as.matrix((column.design)))
colnames(column.design) <- colnames((new.ibm.emp.data1))





#a <- as.data.frame(a)
#colnames(a[1:6]) = colnames(new.ibm.emp.data1[1:6])
#a
#colnames(a)<-colnames(new.ibm.emp.data1)
#column.design[a]= colnames(new.ibm.emp.data1)

#typeof(new.ibm.emp.data1)
  # MFA
#  data('wines2007')
 # demo.mfa.2007 <- mpMFA(wines2007$data, wines2007$table)
#  print(wines2007$data)

#column.design = rep(1, length.out = ncol(new.ibm.emp.data1) )
#column.design[ grep('Year',colnames(new.ibm.emp.data)) ] = 2
#column.design[ grep('Rate',colnames(new.ibm.emp.data)) ] = 3
#column.design = t(as.matrix(column.design))
#colnames(column.design) = colnames(new.ibm.emp.data1)


```


```{r MFA_DESIGN, include=FALSE,echo=FALSE}
#Create an empty list, called DESIGN
DESIGN <- list()


#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$JobLevel$labels <- unique(ibm.emp.data$JobLevel)


#Storing the categorical values in a vector
DESIGN$rows$JobLevel$vec <- ibm.emp.data$JobLevel

#Convert the vector to a matrix
DESIGN$rows$JobLevel$mat <- makeNominalData(as.matrix(DESIGN$rows$JobLevel$vec))

```




```{r MFA_colors_2, include=FALSE,echo=FALSE}

#As our group has 5 levels we specify same number of colors


DESIGN$rows$JobLevel$color_groups <- c("red", "green", "magenta", "blue","yellow")


#plot(c(1:5), pch=15, cex=4, col=DESIGN$rows$JobLevel$color_groups) 

#First, copy the group names
DESIGN$rows$JobLevel$color_observ <- as.matrix(DESIGN$rows$JobLevel$vec)

#Then, for each group, replace the group name with the group's color
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="1")]  <- DESIGN$rows$JobLevel$color_groups[1]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="2")]  <- DESIGN$rows$JobLevel$color_groups[2]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="3")]  <- DESIGN$rows$JobLevel$color_groups[3]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="4")]  <- DESIGN$rows$JobLevel$color_groups[4]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="5")]  <- DESIGN$rows$JobLevel$color_groups[5]

```

## Running MFA
```{r}
#new.ibm.emp.data1 <- subset(new.ibm.emp.data, select = -c(JobLevel))
res.mfa <- mpMFA(new.ibm.emp.data1,as.matrix(column.design),make.columndesign.nominal = TRUE,graphs = FALSE)

```


```{r echo = FALSE}
Eigs <- res.mfa$mexPosition.Data$Table$eigs
tau <- res.mfa$mexPosition.Data$Table$t
Fj <- res.mfa$mexPosition.Data$Table$Q
Cj <- res.mfa$mexPosition.Data$Table$cj

color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(ibm.emp.data))
col4Levels <- data4PCCAR::coloringLevels(rownames(Fj), color4Var)
col4Labels <- col4Levels$color4Levels

varCtr <- data4PCCAR::ctr4Variables(Cj)



absCtrVar <- as.matrix(varCtr) %*% diag(Eigs)
varCtr12 <- (absCtrVar[,1] + absCtrVar[,2]) /(Eigs[1] + Eigs[2])
importantVar <- (varCtr12 >= 1 / length(varCtr12))
col4ImportantVar <- color4Var
col4NS <- 'gray90'
col4ImportantVar[!importantVar] <- col4NS
#Color for importance in  variables
col4Levels.imp <- data4PCCAR::coloringLevels(rownames(Fj),col4ImportantVar)

```

### Scree Plot
```{r}
# A look at the scree
#dev.new()
PlotScree(ev = res.mfa$mexPosition.Data$Table$eigs,
          p.ev = NULL, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F",
          title = "Explained Variance per Dimension - MFA",plotKaiser = TRUE)
MFAScree <- recordPlot()
```
**Scree plot gives us 5 components that are above the Kaiser Line**


### Factor Maps
```{r}
baseMap.i <- createFactorMap(res.mfa$mexPosition.Data$Table$fi, 
                             col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Partial Factor scores by Job Level",display.labels = FALSE,display.points = TRUE)  

x_y_labels <- createxyLabels.gen(
1,2, lambda = res.mfa$mexPosition.Data$Table$eigs,
tau = res.mfa$mexPosition.Data$Table$t
)
#dev.new()

print(baseMap.i$zeMap + x_y_labels)
```

### Loadings
```{r}
baseMap.j.12 <- createFactorMap(Fj, #constraints = constraints.sym,
                             col.points = res.mfa$Plotting.Data$fj.col,col.labels = res.mfa$Plotting.Data$fj.col,
                             display.labels = TRUE,display.points = TRUE,text.cex = 2.5,force=2,cex=1,
                              title = "J Map for Dimension 1 & 2")


#lines4J <- addLines4MCA(Fj, col4Var = col4Levels.imp$color4Variables, size = 1)

Loadings_12 <- baseMap.j.12$zeMap+ x_y_labels

print(Loadings_12)
```


## Inferences
### Means
```{r}
#_____________________________________________________________________
# Inferences

#_____________________________________________________________________
# Mean Map
#  create the map for the means
#  get the means by groups

data_means <- PTCA4CATA::getMeans(res.mfa$mexPosition.Data$Table$fi, ibm.emp.data$JobLevel)
rownames(data_means) = c("JL5","JL4","JL3","JL2","JL1")

# the map
MapGroup    <- PTCA4CATA::createFactorMap(data_means,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)


# The map with observations and group means
a003.Map.I.withMeans <- baseMap.i$zeMap +
                         MapGroup$zeMap_dots + MapGroup$zeMap_text + x_y_labels
#dev.new()
print(a003.Map.I.withMeans)
```


### Partial Factor Scores

Here we calculate the partial factor scores with respect to the 3 tables that we created, to check the inertiaof each.
```{r}

#dim(res.mfa$mexPosition.Data$Table$partial.fi)
data_means1 <- PTCA4CATA::getMeans(res.mfa$mexPosition.Data$Table$partial.fi[1:237,], ibm.emp.data$JobLevel)
data_means2 <- PTCA4CATA::getMeans(res.mfa$mexPosition.Data$Table$partial.fi[238:474,], ibm.emp.data$JobLevel)
data_means3 <- PTCA4CATA::getMeans(res.mfa$mexPosition.Data$Table$partial.fi[475:711,], ibm.emp.data$JobLevel)
rownames(data_means1) = c("Income5","Income4","Income3","Income2","Income1")
rownames(data_means2) = c("Ordinal5","Ordinal4","Ordinal3","Ordinal2","Ordinal1")
rownames(data_means3) = c("Year5","Year4","Year3","Year2","Year1")

# the map
MapGroup1    <- PTCA4CATA::createFactorMap(data_means1,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)

MapGroup2    <- PTCA4CATA::createFactorMap(data_means2,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)

MapGroup3    <- PTCA4CATA::createFactorMap(data_means3,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)

partial_plot <- a003.Map.I.withMeans+MapGroup1$zeMap_dots + MapGroup1$zeMap_text+MapGroup2$zeMap_dots + MapGroup2$zeMap_text+MapGroup3$zeMap_dots + MapGroup3$zeMap_text
print(partial_plot + x_y_labels)

```

### Confidence Intervals
Confidence interval shows no overlap amongst the region of JOb Level except for between Job Level 1 & 2.
```{r}
# Confidence Intervals ----
# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(res.mfa$mexPosition.Data$Table$fi, 
                                 design = DESIGN$rows$JobLevel$color_observ,
                                 niter = 100,
                                 suppressProgressBar = TRUE)


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            p.level = .95
)

a004.Map.I.withCI <-  baseMap.i$zeMap + MapGroup$zeMap_text +  GraphElli

#dev.new()
print(a004.Map.I.withCI + x_y_labels)

```



### Contribution Plot

```{r}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrJ <- res.mfa$mexPosition.Data$Table$cj * sign(res.mfa$mexPosition.Data$Table$Q)

b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(res.mfa$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b003.ctrJ.s.1)


b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                           threshold = 1 / NROW(signed.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(res.mfa$Plotting.Data$fj.col), # we need hex code
                           main = 'Variable Contributions (Signed): Dimension 2',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
#dev.new()
print(b004.ctrJ.s.2)

```
**## Dimension1:**
**Age,Monthly Income,TotalWorkingYears,YearsAtCompany,YearsInCurrentRole are all on the -ve side of Dimension1**
**## Dimension2:**
**WorkLifeBalance & TrainTimeLastYear are on the -ve side of Dimension2**





**WorkLifeBalance and TrainingTimeLastYear are coming at extreme ends of dimension 2, hence are -vely correlated.**

## Conclusion
- In PArtial Factor score map, Year table had more inertia as compared to the other two for higher Job Levels
- SO as Job Level increases, experience increases and it becomes an importatn factor compared to other variables





