---
title: "PLS-C"
#author: "Ajay Attuchirayil Krishnankutty"
#date: "10/26/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
graphics.off()


library(ExPosition)
library(InPosition)
library(xlsx)
library(corrplot)

library(dplyr)
library(ggplot2)
library(PTCA4CATA)
library(data4PCCAR)
library(corrplot)
#install.packages('TExPosition')
library(TExPosition)
library(scales)

file4PLSCfunctions <- 'inferences4PLSC.R'
# Source the function file:
source(file4PLSCfunctions)

```

# PLSC (Partial Least Square Correlation)
PLSC is used when we are trying to find information that is shared between two tables.
We have two latent variables after the analysis, for each of the tables involved.

- Latent Variables : Variables obtained as linear combination of original variables.
- Saliences : Each variables are called saliences and contribute according to their weights towards respective latent       variables.



## Data set: IBM-HR-Emplyee-WithAttrition

The dataset given is regarding IBM employees, available to the HR section of the company to take decisions regarding attrition.

It consists of data of 237 employees (rows), with 32 variables(columns) defining each employee based on position,education,department etc.


```{r data_set,echo=FALSE}
# Dataset ----
ibm.emp.data <- read.xlsx("IBM-HR-Emplyee-WithAttrition.xlsx", sheetName = "Sheet1",header=TRUE)
rownames(ibm.emp.data) <- ibm.emp.data$Subj
ibm.emp.data <- ibm.emp.data[-c(1)]
cols <- c('Age','Gender','Education','MonthlyIncome','TotalWorkingYears','JobSatisfaction')
head(ibm.emp.data[cols])

```

### Splitting the data into 3 tables, 2 for the correaltion operation, 3rd for the deisgn
```{r}

ibm.year <- ibm.emp.data[10:13]
ibm.income <- ibm.emp.data[20:28]
ibm.income <- ibm.income[-c(5)]
colnames(ibm.income)
colnames(ibm.year)
ibm.design <- ibm.emp.data[,18]
raw.data <- ibm.emp.data[-c(10:13,20:28)]

```

### Take smal sample out of the tables
```{r}

year <- ibm.year#[1:nN,]
income <- ibm.income#[1:nN,]
sample.design <- ibm.design#[1:nN]

```

### Correlation plot amongst the sample
```{r}
## heatmap
corrplot::corrplot(cor(year,income))
```


### Creating DESIGN variables.

#### 1st Design Variable : Job Level

```{r DESIGN,echo=FALSE}
#Create an empty list, called DESIGN
DESIGN <- list()
#We take all the unique values possible in the specified column if the dataframe
DESIGN$rows$JobLevel$labels <- unique(ibm.emp.data$JobLevel)
#Storing the categorical values in a vector
DESIGN$rows$JobLevel$vec <- ibm.emp.data$JobLevel
#Convert the vector to a matrix
DESIGN$rows$JobLevel$mat <- makeNominalData(as.matrix(DESIGN$rows$JobLevel$vec))

```



```{r colors_2,echo=FALSE}

#As our group has 5 levels we specify same number of colors
DESIGN$rows$JobLevel$color_groups <- c("red", "green", "magenta", "blue","yellow")
#plot(c(1:5), pch=15, cex=4, col=DESIGN$rows$JobLevel$color_groups) 

#First, copy the group names
DESIGN$rows$JobLevel$color_observ <- as.matrix(DESIGN$rows$JobLevel$vec)

#Then, for each group, replace the group name with the group's color
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="1")]  <- DESIGN$rows$JobLevel$color_groups[1]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="2")]  <- DESIGN$rows$JobLevel$color_groups[2]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="3")]  <- DESIGN$rows$JobLevel$color_groups[3]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="4")]  <- DESIGN$rows$JobLevel$color_groups[4]
DESIGN$rows$JobLevel$color_observ[which(DESIGN$rows$JobLevel$vec=="5")]  <- DESIGN$rows$JobLevel$color_groups[5]

```




## Running PLSC
```{r}
resPLSC <- tepPLS(year,income,DESIGN = factor(sample.design),graphs = FALSE)

#resPCA.inf <- epPCA.inference.battery(year,income,DESIGN = sample.design,graphs = FALSE)

Fi   <- resPLSC$TExPosition.Data$fi
Fj   <- resPLSC$TExPosition.Data$fj
constraints.factor <- minmaxHelper(mat1 = Fi, mat2  = Fj)

Eigs <- resPLSC$TExPosition.Data$eigs
tau <- resPLSC$TExPosition.Data$t

Cj <- resPLSC$TExPosition.Data$cj

label2PLSC <- createxyLabels.gen(x_axis = 1, y_axis = 2,
              lambda = Eigs,tau =tau)

Lx <- resPLSC$TExPosition.Data$lx 
Ly <- resPLSC$TExPosition.Data$ly
constraints.latent <- minmaxHelper(mat1 = Lx, mat2  = Ly)
```

### Permutation Test
```{r include=FALSE}
#--------------------------------------------------------------------
# First: Go for a permutation test
#
resPerm4PLSC <- perm4PLSC(year, # First Data matrix 
                          income, # Second Data matrix
                          nIter = 1000 # How mny iterations
                          )
# to see what results we have
print(resPerm4PLSC)

```




### Scree Plot of Eigen values 
```{r}
# Eigenv
EigVal <- resPLSC$TExPosition.Data$eigs 
# How many eigen-value do we have here
nL <- min(ncol(year),ncol(income))

PlotScree(ev = EigVal, 
       p.ev = resPerm4PLSC$pEigenvalues,
       title = "Scree Plot with Permutation Test",
       plotKaiser = TRUE
          )
a001a.screePlot <- recordPlot()

```


### Salience - Fi
```{r}
salience.i <- createFactorMap(Fi, constraints = constraints.factor,
                             col.points = resPLSC$Plotting.Data$fi.col ,
                             col.labels = 'darkorchid',title = "Saliences - Fi",display.labels = TRUE)  

#dev.new()
#hist(Fi)
print(salience.i$zeMap  + label2PLSC)
```



### Salience - Fj
```{r}
salience.j <- createFactorMap(Fj, constraints = constraints.factor,
                             col.points = resPLSC$Plotting.Data$fj.col ,
                             col.labels = 'darkorchid',title = "Saliences - Fj",display.labels = TRUE)  

#dev.new()

print(salience.j$zeMap  + label2PLSC)

```


---- Design Variable : Job Level
**We will plot Saliences and Factor scores and bootstrap ratios side by side, to show the effect on one table's variable based on the pattern in the second table**

### LX1 vs Ly1
```{r}
Lx1 <- resPLSC$TExPosition.Data$lx[,1]

Ly1 <- resPLSC$TExPosition.Data$ly[,1]

df <- cbind(Lx1,Ly1)


baseMap.i <- createFactorMap(df,# constraints = constraints.sym,
                             col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Lx1 & Ly1 by Job Level",display.labels = FALSE)  

print(baseMap.i$zeMap_background +baseMap.i$zeMap_dots + label2PLSC)


```

## Inferences

### Means between Lx1 & Ly1

```{r}


col4data0 <- resPLSC$Plotting.Data$fii.col
col4Means0 <- unique(col4data0)


data_means <- PTCA4CATA::getMeans(df,ibm.emp.data$JobLevel)
# a vector of color for the means
rownames(data_means) = c("JL5","JL4","JL3","JL2","JL1")


MapGroup    <- PTCA4CATA::createFactorMap(data_means,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = DESIGN$rows$JobLevel$color_groups,
                            cex = 7,  # size of the dot (bigger)
                            display.labels = TRUE,
                            col.labels = DESIGN$rows$JobLevel$color_groups,
                            text.cex = 6)


# The map with observations and group means
a003.Map.I.withMeans <- baseMap.i$zeMap +MapGroup$zeMap_dots + MapGroup$zeMap_text
#a003.Map.I.withMeans <- baseMap.i$zeMap +MapGroup$zeMap_dots + MapGroup$zeMap_text
                         
#dev.new()
print(a003.Map.I.withMeans)
```


```{r include=FALSE}
BootCube <- PTCA4CATA::Boot4Mean(df, 
                                 design = ibm.emp.data$JobLevel,
                                 niter = 100,
                                 suppressProgressBar = TRUE)
```


### Confidence Intervals between Lx1 & Ly1

```{r eval=FALSE,include=FALSE}
# Confidence Intervals ----
# Bootstrap for CI:


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            p.level = .95
)


a004.Map.I.withCI <- baseMap.i$zeMap + MapGroup$zeMap_text +  GraphElli


dev.new()
print(a004.Map.I.withCI)

```


### Tolerance Intervals

```{r}
# Tolerance Intervals ----

GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(df,
                            design = ibm.emp.data$JobLevel,
                            # line below is needed
                            names.of.factors =  c("Dim1","Dim2"), # needed 
                            #col = rownames(BootCube$BootCube[,c(1,2),1]),
                            col = DESIGN$rows$JobLevel$color_groups,
                            line.size = .50, 
                            line.type = 3,
                            alpha.ellipse = .2,
                            alpha.line    = .4,
                            p.level       = .75)


a005.Map.I.withTIHull <-baseMap.i$zeMap  +
                            GraphTI.Hull + MapGroup$zeMap_dots +
                           MapGroup$zeMap_text + MapGroup$zeMap_dots


#dev.new()
print(a005.Map.I.withTIHull)

```


### Lx2 vs Ly2

```{r}
Lx2 <- resPLSC$TExPosition.Data$lx[,2]
Ly2 <- resPLSC$TExPosition.Data$lx[,2]
df1 <- as.data.frame(cbind(Lx2,Ly2))
#ggplot(NULL, aes(x=Lx2,y=Ly2)) + geom_point(aes(colour = ibm.emp.data$JobLevel))


baseMap1 <- createFactorMap(df1,# constraints = constraints.sym,
                             col.points = DESIGN$rows$JobLevel$color_observ,
                             col.labels = 'darkorchid',title = "Lx2 & Ly2 by Job Level",display.labels = FALSE)  


print(baseMap1$zeMap)


```



### Means between Lx2 & Ly2

```{r}
#_____________________________________________________________________
# Inferences

#_____________________________________________________________________
# Mean Map
#  create the map for the means
#  get the means by groups

data_means <- PTCA4CATA::getMeans(df1,ibm.emp.data$JobLevel)
rownames(data_means) =  c("JL5","JL4","JL3","JL2","JL1")
# a vector of color for the means
col4Data <- DESIGN$rows$JobLevel$color_groups
#col4Means <- c('#F8766D','#A3A500','#00BF7D','#00B0F6','#E76BF3')
# the map
MapGroup    <- PTCA4CATA::createFactorMap(data_means,
                            # use the constraint from the main map
                            #constraints = baseMap.i$constraints,
                            col.points = col4Data,
                            cex = 7,  # size of the dot (bigger)
                            col.labels =col4Data,
                            text.cex = 6)


# The map with observations and group means
#a003.Map.Y.withMeans <- baseMap1.i$zeMap +MapGroup$zeMap_dots + MapGroup$zeMap_text
a003.Map.Y.withMeans <- baseMap1$zeMap +MapGroup$zeMap_dots + MapGroup$zeMap_text + label2PLSC
                         
#dev.new()
print(a003.Map.Y.withMeans)
```


**Lx2 & Ly2 showed perfect correlation**



### Confidence Intervals  between Lx2 & Ly2
```{r include=FALSE,eval=FALSE}
# Confidence Intervals ----
# Bootstrap for CI:
BootCube <- PTCA4CATA::Boot4Mean(df1, 
                                 design = DESIGN$rows$JobLevel$color_observ,
                                 niter = 100,
                                 suppressProgressBar = TRUE)

#_____________________________________________________________________
# Create the ellipses
#_____________________________________________________________________
# Create Confidence Interval Plots
# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            p.level = .95
)
#_____________________________________________________________________
# create the I-map with Observations, means and confidence intervals
#

#z1 <- baseMap + GraphElli +MapGroup$zeMap_dots + MapGroup$zeMap_text

a004.Map.Y.withCI <- baseMap1 + MapGroup$zeMap_text +  GraphElli
#_____________________________________________________________________
# plot it!
#dev.new()
print(a004.Map.Y.withCI)

```


### Tolerance Intervals

```{r}
# Tolerance Intervals ----
# use function MakeToleranceIntervals from package PTCA4CATA
GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(df1,
                            design = as.factor(sample.design),
                            # line below is needed
                            names.of.factors =  c("Dim1","Dim2"), # needed 
                            col = rownames(BootCube$BootCube[,c(1,2),1]),
                            line.size = .50, 
                            line.type = 3,
                            alpha.ellipse = .2,
                            alpha.line    = .4,
                            p.level       = .75)
#_____________________________________________________________________
# Create the map:
a005.Map.Y.withTIHull <-a003.Map.Y.withMeans  +
                            GraphTI.Hull #+ MapGroup$zeMap_dots +
                           #MapGroup$zeMap_text + MapGroup$zeMap_dots + label2PLSC


#z4 <- 

#_____________________________________________________________________
# plot it
#dev.new()
print(a005.Map.Y.withTIHull)

```







### Contribution Bars
```{r echo=FALSE}
#_____________________________________________________________________
# Contribution Plots ----
# get the Contributions and make a plot.
#_____________________________________________________________________
# Here we look only at the (signed) contributions for the variables
# compute teh signed contributions
signed.ctrI <- resPLSC$TExPosition.Data$ci * sign(resPLSC$TExPosition.Data$fi)

b003.ctr.fi.1 <- PrettyBarPlot2(signed.ctrI[,1],
                         threshold = 1 / NROW(signed.ctrI),
                         font.size = 5,
                         color4bar = gplots::col2hex(resPLSC$Plotting.Data$fi.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrI), 1.2*max(signed.ctrI)),horizontal=FALSE
)
#dev.new()
print(b003.ctr.fi.1)


signed.ctrJ <- resPLSC$TExPosition.Data$cj * sign(resPLSC$TExPosition.Data$fj)

b003.ctr.fj.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(resPLSC$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 1',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ)),horizontal=FALSE
)
#dev.new()
print(b003.ctr.fj.1)



#signed.ctrI <- resPLSC$TExPosition.Data$ci * sign(resPLSC$TExPosition.Data$fi)

b003.ctr.fi.2 <- PrettyBarPlot2(signed.ctrI[,2],
                         threshold = 1 / NROW(signed.ctrI),
                         font.size = 5,
                         color4bar = gplots::col2hex(resPLSC$Plotting.Data$fi.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 2',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrI), 1.2*max(signed.ctrI)),horizontal=FALSE
)
#dev.new()
print(b003.ctr.fi.2)


#signed.ctrJ <- resPLSC$TExPosition.Data$cj * sign(resPLSC$TExPosition.Data$fj)

b003.ctr.fj.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(resPLSC$Plotting.Data$fj.col), # we need hex code
                         main = 'Variable Contributions (Signed): Dimension 2',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ)),horizontal=FALSE
)
#dev.new()
print(b003.ctr.fj.2)


```

### Bootstrap Test
```{r include=FALSE}
#--------------------------------------------------------------------
# Second: Compute the Bootstrap ratios
#
#' 
resBoot4PLSC <- Boot4PLSC(year, # First Data matrix 
                          income, # Second Data matrix
                          nIter = 1000, # How many iterations
                      Fi = resPLSC$TExPosition.Data$fi,
                      Fj = resPLSC$TExPosition.Data$fj,
                      nf2keep = 3,
                      critical.value = 2,
                      # To be implemented later
                      # has no effect currently
                      alphaLevel = .05)
#
# to see what results we have
#print(resBoot4PLSC)
# Exercise: Plot the Bootstrap ratios
# along with their critical value. 
# Question what do we conclude?
#--------------------------------------------------------------------



```

### Bootstrap Ratios for first table
```{r}
laDim = 1
ba001.BR1 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.i[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resPLSC$Plotting.Data$fi.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',horizontal=FALSE
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR1)


laDim = 2
ba001.BR2 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.i[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resPLSC$Plotting.Data$fi.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',horizontal=FALSE
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR2)




```

**1st Table: Monthly Income on the -ve side of DImension 1**

### Bootstrap Ratios for second table
```{r}
laDim = 1
ba001.BR1 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.j[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resPLSC$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',horizontal=FALSE
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR1)


laDim = 2
ba001.BR2 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.j[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(resPLSC$Plotting.Data$fj.col), # we need hex code
                   main = paste0(
                     'Bootstrap ratio for Dimension:',laDim),
                  ylab = 'Bootstrap ratios',horizontal=FALSE
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
#dev.new()
print(ba001.BR2)




```

**2nd Table: Year realted variables are on the -ve side of DImension 1, and pecent salry hike is on +ve side of Dimension 1**


## Conclusion
- Saliences Fi when considered with factor score plot, showed higher Job Levels coincided with higher monthly income.  
- Saliences Fj when considered with factor score plot, showed similar trend as above with all the year related 	variables.  	



```{r eval=FALSE,include=FALSE}
listSaved <- saveGraph2pptx(
file2Save.pptx = '[Group2_Ajay]PLSC-IBM-HR-Employee-WithAttrition.pptx', 
title = "PLSC", 
addGraphNames = TRUE)
```



